---
title: "Diet composition analysis"
author: "Nathan Brouwer"
date: "February 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

### Load trait dta
```{r}
load("./data/trait_dat.RData")
```

### Load libraries

```{r}
source('~/1_R/git/git-aviary/DRmencia/R/load_libraries.R')
load_libraries()
library(scales) #for stacked bargraph

library(grid)
library(wrapr)
library(reshape2)
```

### Data structure

Number of individuals in each diet category in a sep. row

      site      year year.num site.age diet.C diet.F diet.G diet.I
1 La Cueva 2003-2004     2003        2      2      2     43     49
2 La Cueva 2004-2005     2004        3      2     12     70    103
3 La Cueva 2005-2006     2005        4      5      1     43     52
4 La Cueva 2006-2007     2006        5      5      2     34     85
5 La Cueva 2007-2008     2007        6      4      6     17     47
6 La Caoba 2003-2004     2003        5      5      1     32     85
```{r}
head(trait_dat)[,c(1:4,10:13)]
```





### Melt to  long

```{r}
id.cols <- c("site","year",
             "year.num","site.age",
             "site.age.init", "site.age.cent")

#wrapr::qc() quotes text
diet.cols <- qc(diet.C,diet.F,diet.G,diet.I,diet.N,diet.O)

diet.raw.melt <- melt(trait_dat,
                     id.vars = id.cols,
                     measure.vars = diet.cols,
                     variable.name = "Diet category")

```


### Clean
```{r}
names(diet.raw.melt) <- gsub("value","count",names(diet.raw.melt))

diet.raw.melt$Diet <- gsub("diet.","",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("C","Carnivore",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("F","Frugivore",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("G","Granivore",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("N","Nectarivore",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("I","Insectivore",diet.raw.melt$Diet)
diet.raw.melt$Diet <- gsub("O","Omnivore",diet.raw.melt$Diet)
```

## Plot

### Plot stacked bargraph using ggplot
Uses scales library

```{r}
gg.diet.stacked <- ggplot(diet.raw.melt, aes(x = site, 
                                             y = count,
                                             fill = Diet)) + 
    geom_bar(position = "fill",stat = "identity") +
    # or:
    # geom_bar(position = position_fill(), stat = "identity") 
    scale_y_continuous(labels = percent_format()) +
  xlab("Site") +
  ylab("Percent (%)") 
```


grobs uses grid

https://stackoverflow.com/questions/12409960/ggplot2-annotate-outside-of-plot


```{r}
library(RColorBrewer)

gg.diet.stacked+ annotate("text", 
                          x = 1, 
                          y = 0.125, 
                          label = "Omnivore",
                          col = "white", 
                          size = 5) +
  annotate("text", x = 1, y = 0.3725, label = "Nectarivore",col = "white", 
           size = 5) +
  annotate("text", x = 1, y = 0.64575, label = "Insectivore",col = "white", 
           size = 5) +
  annotate("text", x = 1, y = 0.875, label = "Granivore",col = "white", 
           size = 5) +
  annotate("text", x = 4, y = 0.982775, label = "Carnivore",col = "white", 
           size = 5) 
  
  


```

## Plot stacked bargraph using barplot()

https://www.r-graph-gallery.com/212-stacked-percent-plot/
https://www.r-graph-gallery.com/212-stacked-percent-plot/

```{r}
i.yr.2004 <- which(trait_dat$year == "2003-2004" | trait_dat$year == "1996-1997")
dat <- trait_dat[i.yr.2004,c("diet.C","diet.F","diet.G","diet.I","diet.N","diet.O")]
sites. <- trait_dat[i.yr.2004,c("site")]

#Transform this data in %
dat_percentage=apply(t(dat), 2, function(x){x*100/sum(x,na.rm=T)})

# Make a stacked barplot--> it will be in %!
#col=coul 
coul = brewer.pal(3, "Pastel2") 
barplot(dat_percentage, 
        border="black", 
        xlab="group", 
        col = c(1,2,3,4,5),
        names.arg = sites.,
        density=c(50,100,50,100,50),
        angle=c(0,45,90,11),
        legend.text = TRUE,
        args.legend = list(cex =1.75),
        xlim=c(0, ncol(dat_percentage) + 3))


```


### NMDS

#### Fit NMDS
```{r}
j.use <- grep("diet",names(trait_dat))
names(trait_dat) <- gsub("diet.","",names(trait_dat))
names(trait_dat) <- gsub("C","Carnivore",names(trait_dat))
names(trait_dat)<- gsub("F","Frugivore",names(trait_dat))
names(trait_dat) <- gsub("G","Granivore",names(trait_dat))
names(trait_dat) <- gsub("N","Nectarivore",names(trait_dat))
names(trait_dat) <- gsub("I","Insectivore",names(trait_dat))
names(trait_dat) <- gsub("O","Omnivore",names(trait_dat))

diet_NMDS <- metaMDS(trait_dat[,j.use])

```

#### Stress plot

```{r}
stressplot(diet_NMDS)
```


#### Plot output
```{r}
ordiplot(diet_NMDS)
orditorp(diet_NMDS,display="species")
#orditorp(diet_NMDS,display="sites")
ordiellipse(diet_NMDS,
         groups=trait_dat$site,
         label=T,
         kind = "se")
```



## Analysis of pecetages

## Convert raw counts to compositions

```{r}
library(compositions)
load("./data/trait_dat.RData")
trait_dat_comp <- trait_dat

trait_dat_comp[,diet.cols] <- clo(trait_dat[,diet.cols])

```


```{r}
diet.comp.melt <- melt(trait_dat_comp,
                     id.vars = id.cols,
                     measure.vars = diet.cols,
                     variable.name = "diet")
names(diet.comp.melt) <- gsub("value","Percent",names(diet.comp.melt))

```


### Model with regression

Model one of the components with regression.  Use logit transformation
Model just 2ndary forest
```{r}
i.diet.O.no.Aceit <- which(diet.comp.melt$diet == "diet.O" & 
                diet.comp.melt$site != "Aceitillar")

diet.comp.0 <- lmer(boot::logit((Percent)) ~ 1 + 
                (1|site) +
                (1|year),
              data = diet.comp.melt[i.diet.O.no.Aceit,],
              control = lmerControl(optimizer = "Nelder_Mead"),
              REML = FALSE)

diet.comp.age <- update(diet.comp.0, . ~ . + site.age.cent)
diet.comp.age2 <- update(diet.comp.0, . ~ . + 
                               site.age.cent +
                               I(site.age.cent^2))


```


### Inference

Look at rank of all 3 models
```{r}
ICtab(diet.comp.0,
       diet.comp.age,
       diet.comp.age2,
      type = "AICc")
```


```{r}
summary(diet.comp.age)
summary(diet.comp.age2)
```


```{r}
anova(diet.comp.0,
      diet.comp.age)

anova(diet.comp.age,
      diet.comp.age2)
```


Confidence intervals

```{r}
confint.merMod(diet.comp.age,method = "boot")
```





## 1-way ANOVA


### Remove overlap

```{r}
i.cueva.diet.drop <- which(diet.comp.melt$site == "La Cueva" & 
                      diet.comp.melt$site.age > 4) #keep age 2, 3, 4
i.caoba.diet.drop <- which(diet.comp.melt$site == "La Caoba" & 
                      diet.comp.melt$site.age < 7) #keep age 7, 8, 9

diet.comp.melt_2 <- diet.comp.melt[-c(i.cueva.diet.drop,i.caoba.diet.drop), ]
```


Model diet.O
```{r}
i.diet.O2 <- which(diet.comp.melt_2$diet == "diet.O")
diet.comp.ANOVA.means <- blmer(boot::logit((Percent)) ~ -1 + site +
                #(1|site) +
                (1|year),
              data = diet.comp.melt_2[i.diet.O2,])

diet.comp.ANOVA.means2 <- blmer(((Percent)) ~ -1 + site +
                #(1|site) +
                (1|year),
              data = diet.comp.melt_2[i.diet.O2,])
```

```{r}
summary(diet.comp.ANOVA.means)
plot(fixef(diet.comp.ANOVA.means) ~ c(1,2,3,4,5))
```

## CIs
                    2.5 %      97.5 %
.sig01          0.22125648  0.70754225
.sigma          0.09057722  0.23605816
siteLa Cueva   -1.61278756 -0.76777985
siteLa Caoba   -1.43047644 -0.57911794
siteMorelia    -1.38816835 -0.64996773
siteEl Corral  -1.14454409 -0.35057166
siteAceitillar -0.60883944  0.07904538

siteLa Cueva   siteLa Caoba    siteMorelia  siteEl Corral siteAceitillar 
    -1.1819153     -1.0114089     -1.0263291     -0.7459612     -0.2738489 
    
```{r}
hab.comp.CIs <- confint(diet.comp.ANOVA.means, method = "boot")
```

```{r}
invlogit(c(-1.61278756,-0.76777985))
invlogit(c(-0.60883944,0.07904538))

invlogit(fixef(diet.comp.ANOVA.means))
```



## Contrast modeling

Create contrast matrix
```{r}
contrst.mat.traits <- rbind("linear-traits" = c(-2,-1, 0, 1, 2),
                     "quad-traits"   = c(+2,-1,-2,-1,+2))


   
```


Run multcomp
```{r}
mult.comp.working.diet <- glht(hab.comp.ANOVA.means       
                          ,linfct =contrst.mat.traits
                          ,alternative = c("two.sided"))


mult.comp.diet.summary <- summary(mult.comp.working.diet                            ,test = adjusted(type = "none"))
```



### Save

```{r}
save.image(file = "./models/WRKSPC_diet_composition.RData")
```


```{r}

```

