---
title: "Analysis of preffered habitats"
author: "Nathan Brouwer"
date: "February 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

### Load trait data
```{r}
load("./data/trait_dat.RData")
```

### Load libraries
```{r}
source('~/1_R/git/git-aviary/DRmencia/R/load_libraries.R')
load_libraries()

#Load compositions library
## for compositional data analysis
library(compositions)
```



### Data structure

Columns for number of individuals in each site-year combo that were of a given habitat preference group (DF,EF,SF)

```{r}
head(trait_dat)[,c(1:3,7:16)]

dim(trait_dat)
```
      site      year year.num hab.DF hab.EF hab.SF diet.C diet.F diet.G diet.I
1 La Cueva 2003-2004     2003     14     70     85      2      2     43     49
2 La Cueva 2004-2005     2004     46    194    131      2     12     70    103
3 La Cueva 2005-2006     2005     17     96     72      5      1     43     52
4 La Cueva 2006-2007     2006     19    133     75      5      2     34     85
5 La Cueva 2007-2008     2007     20     77     28      4      6     17     47
6 La Caoba 2003-2004     2003     10    127    105      5      1     32     85


```{r}
with(trait_dat,
     table(site,year))
```





### Transform habitat data
```{r}
id.cols <- c("site","year",
             "year.num","site.age",
             "site.age.init", "site.age.cent")
hab.cols <- c("hab.DF","hab.EF","hab.SF")

## create closure
hab.clo <- clo(trait_dat[,hab.cols])


## centered log ratio
hab.clr <- clr(hab.clo)

#isometric log ratio
hab.ilr <- ilr(hab.clo)




# 
# hab.clr <- cbind(hab.clr,
#                  trait_dat[,id.cols])
# hab.ilr <- cbind(hab.ilr,
#                  trait_dat[,id.cols])
```


## Model compositional data

### Model w/MANOVA

This is exploratory; treats all years as independent
```{r}
m.hab.clr          <- manova(hab.clr ~ trait_dat$site)
m.hab.ilr.site.age <- manova(hab.ilr ~ trait_dat$site.age)
m.hab.ilr.site     <- manova(hab.ilr ~ trait_dat$site)
```


```{r}
summary(m.hab.ilr.site.age)
summary(m.hab.ilr.site)
```



### Model with permanova

Note: can't really use "strata" b/c there is only 1 measurement per year for Aceitillar

http://r-sig-ecology.471788.n2.nabble.com/Measurement-distance-for-proportion-data-td7578891.html

"I would also suggest to give a try to the Aitchison distance. To do so, you can use the compositions package. You transform the proportions to centered log-ratios or isometric log-ratios (clr and ilr functions, respectively), then compute the Euclidean distance through transformed data - both transformations should return the same distances. 


comp = acomp(AnimalVegetation[,1:4]) # proportions closed between 0 and 1 

bal = ilr(comp) # isometric log-ratios 

dist = vegdist(bal, method="euclidean") # Aitchison dissimilarity matrix 
mod = betadisper(dist, region) 
mod 
plot(mod) 
adonis(dist ~ region) 


```{r}
acomp. <- acomp(trait_dat[,hab.cols])# proportions closed between 0 and 1 
acomp.ilr <- ilr(acomp.)# isometric log-ratios 
acomp.ilr.dist = vegdist(acomp.ilr, method="euclidean")# Aitchison dissimilarity matrix 
#betadisper() homogen of dispersion
acomp.ilr.dist.adonis <- adonis(acomp.ilr.dist ~  trait_dat$site,
                                strata = trait_dat$year) 
acomp.ilr.dist.adonis
summary(acomp.ilr.dist.adonis)
```



## Reshape data

### Melt to long

```{r}
library(reshape2)
#add "closed" data to ID columns
hab.clo2 <- cbind(hab.clo,
                  trait_dat[,id.cols])

hab.clo.melt <- melt(hab.clo2,
                     id.vars = id.cols,
                     measure.vars = hab.cols,
                     variable.name = "Habitat")
names(hab.clo.melt) <- gsub("value","Percent",names(hab.clo.melt))

```



### Model with regression

Model one of the components with regression.  Use logit transformation
Model just 2ndary forest
```{r}
i.SF.no.Aceit <- which(hab.clo.melt$Habitat == "hab.SF" & 
                hab.clo.melt$site != "Aceitillar")

hab.clo.0 <- blmer(boot::logit((Percent)) ~ 1 + 
                (1|site) +
                (1|year),
              data = hab.clo.melt[i.SF.no.Aceit,])

hab.clo.site.age <- update(hab.clo.0, . ~ . + site.age.cent)
hab.clo.site.age2 <- update(hab.clo.0, . ~ . + site.age.cent + I(site.age.cent^2))

ICtab(hab.clo.0,
       hab.clo.site.age,
       hab.clo.site.age2,
      type = "AICc")
```


```{r}
summary(hab.clo.site.age)
summary(hab.clo.site.age2)
```


```{r}
anova(hab.clo.0,
      hab.clo.site.age)

anova(hab.clo.site.age2,
      hab.clo.site.age)
```


```{r}
confint.merMod(hab.clo.site.age2,method = "boot")
```




## Plot
```{r}
i.EF <- which(hab.clo.melt$Habitat == "hab.EF")
i.SF <- which(hab.clo.melt$Habitat == "hab.SF")

ef.lims <- c(0.4,0.9)
sf.lims <- c(0.0,0.6)
gg.ef <- ggplot(data = hab.clo.melt[i.EF,],
       aes(y = Percent,
           x = site.age,
           group = Habitat,
           color = site)) +
  geom_point() +
    ylim(ef.lims) +
  ylab("Proportion") +
  #facet_wrap(~Habitat,nrow = 2) + 
  theme(legend.position="none")+
  theme(#strip.background = element_blank(), 
        strip.text = element_blank()) +
  xlab("")

gg.sf <- ggplot(data = hab.clo.melt[i.SF,],
       aes(y = Percent,
           x = site.age,
           group = Habitat,
           color = site)) +
  geom_point() +
  ylim(sf.lims)+
    ylab("Proportion") +
  #facet_wrap(~Habitat,nrow = 2) + 
  theme(legend.position="none")+
  theme(#strip.background = element_blank(), 
        strip.text = element_blank())

i.ef.Aceit <- which(hab.clo.melt$Habitat == "hab.EF" &
                       hab.clo.melt$site == "Aceitillar")
i.sf.Aceit <- which(hab.clo.melt$Habitat == "hab.SF" &
                       hab.clo.melt$site == "Aceitillar")
aceit.ef <- ggplot(data = hab.clo.melt[i.ef.Aceit,],
       aes(y = Percent,
           x = site,
           group = Habitat,
           color = site)) +
  geom_point() +
  xlab("") + ylab("") +
  ylim(ef.lims) +
  #facet_wrap(~Habitat,nrow = 2)+ 
  theme(legend.position="none") +
  theme(#strip.background = element_blank(), 
        strip.text = element_blank()) +

  theme(axis.title.y=element_blank(),
        axis.line.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  xlab("")



aceit.sf <- ggplot(data = hab.clo.melt[i.sf.Aceit,],
       aes(y = Percent,
           x = site,
           group = Habitat,
           color = site)) +
  geom_point() +
   ylim(sf.lims)+
  xlab("") + ylab("") +
  #facet_wrap(~Habitat,nrow = 2)+ 
  theme(legend.position="none") +
  theme(#strip.background = element_blank(), 
        strip.text = element_blank()) +

  theme(axis.title.y=element_blank(),
        axis.line.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())



cowplot::plot_grid(#gg.ef,aceit.ef,
                   gg.sf,aceit.sf,
                   rel_widths = c(3,1),
                   labels = c(#"Evergreen forest species","",
                              "Secondary forest species",""))
```



```{r}

```


## 1-way ANOVA


### REmove overlap

```{r}
i.cueva.drop <- which(hab.clo.melt$site == "La Cueva" & 
                      hab.clo.melt$site.age > 4) #keep age 2, 3, 4
i.caoba.drop <- which(hab.clo.melt$site == "La Caoba" & 
                      hab.clo.melt$site.age < 7) #keep age 7, 8, 9

hab.clo.melt_2 <- hab.clo.melt[-c(i.cueva.drop,i.caoba.drop), ]
```


Model secondary forest
```{r}
i.SF <- which(hab.clo.melt_2$Habitat == "hab.SF")
hab.clo.ANOVA.means <- blmer(boot::logit((Percent)) ~ -1 + site +
                #(1|site) +
                (1|year),
              data = hab.clo.melt_2[i.SF,])
```

```{r}
summary(hab.clo.ANOVA.means)
```

Model Evergreen forest
```{r}
i.EF <- which(hab.clo.melt_2$Habitat == "hab.EF")
hab.clo.ANOVA.means.EF <- blmer(boot::logit((Percent)) ~ -1 + site +
                #(1|site) +
                (1|year),
              data = hab.clo.melt_2[i.EF,])

summary(hab.clo.ANOVA.means.EF)
```



## Contrast modeling

Create contrast matrix
```{r}
contrst.mat.traits <- rbind("linear-traits" = c(-2,-1, 0, 1, 2),
                     "quad-traits"   = c(+2,-1,-2,-1,+2))


   
```


Run multcomp
```{r}
mult.comp.working.hab <- glht(hab.clo.ANOVA.means       
                          ,linfct =contrst.mat.traits
                          ,alternative = c("two.sided"))


mult.comp.hab.summary <- summary(mult.comp.working.hab
                             ,test = adjusted(type = "none"))
```



### Save

```{r}
save.image(file = "./models/WRKSPC_diet_composition.RData")
```
