---
title: "Analysis: age ratio - 1-way ANOVA with trend test"
author: "Nathan Brouwer"
date: "January 17, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use of 1-way ANOVA style models to determine if age ratios vary accross entire chronosequence of abandoned pastures, including forest reference site.  Trend tests using multcomp used to assess directional change.

This only models the mean trend across all species, NOT species-specific trends.  This would require seperate ANOVAs and trend tests for each species.

## Preliminaries

### Libraries

Function that calls all libraries needed for analysis
```{r}
source("./R/load_libraries.R")
load_libraries()
```


### Load data

* Data has just focal species; rarities removed
* Within-year repatures removed

```{r}
load(file = "./data/ann_caps4age_ratio.RData")
```



### Remove overlap of Youngest sites

The 2 youngest sites both have ages that equal 5 and 6 years: La Cueva's last 2 years, La Caoaba's 1st 2 years.  Using "site" as a factor variable in ANOVA will confound site and age.  This code subsets the data for these sites so thatall Cueva data is younger than Caoba data.

```{r}
with(ann_caps4age_ratio, table(site,site.age))

```


```{r}
i.cueva.drop <- which(ann_caps4age_ratio$site == "La Cueva" & 
                      ann_caps4age_ratio$site.age > 4) #keep age 2, 3, 4
i.caoba.drop <- which(ann_caps4age_ratio$site == "La Caoba" & 
                      ann_caps4age_ratio$site.age < 7) #keep age 7, 8, 9

ann_caps4age_ratio_2 <- ann_caps4age_ratio[-c(i.cueva.drop,i.caoba.drop), ]
```



## Model 

### Base ANOVA model - age ratio

Note that this model uses all species, if even they are rare at Aceitillar; for the Aceitillar only analysis really rare spp were removed.

```{r}
m.ANOVA.0 <- bglmer(as.numeric(age.AHY.01)-1 ~ 1 +
               (1|band) +
               (1|year) +
               (1|site) +
               (1|spp.code) ,#+
               #(1|site:spp.code) #this would be ideal but prob overkill
              data = ann_caps4age_ratio_2[,],
              family = binomial,
              glmerControl(optimizer = "Nelder_Mead")
              )
```



### Run fixed effects
```{r}
m.ANOVA.site          <- update(m.ANOVA.0, . ~ . + site)
m.ANOVA.status.focals <- update(m.ANOVA.0, . ~ . +         status.focals)
m.ANOVA.add           <- update(m.ANOVA.0, . ~ . +  site + status.focals)
m.ANOVA.full          <- update(m.ANOVA.0, . ~ . +       site*status.focals)

#calcualte mean values for each year
m.ANOVA.means         <- update(m.ANOVA.0, . ~ . +  -1 + site:status.focals)
```




### Inference
```{r}
ICtab(m.ANOVA.0,
       m.ANOVA.site,
       m.ANOVA.status.focals,
       m.ANOVA.add,
       m.ANOVA.full,
        type = "AICc")
```



## Contrast modeling

Create contrast matrix
```{r}
contrst.mat <- rbind("Mig-linear" = c(-2,-1, 0, 1, 2, 0, 0, 0, 0, 0),
                     "Mig-quad"   = c(+2,-1,-2,-1,+2, 0, 0, 0, 0, 0),
                     "Mig-brk"    = c(-2, 0, 1, 2, -1, 0, 0, 0, 0, 0),
                     "Res-linear" = c( 0, 0, 0, 0, 0,-2,-1, 0, 1, 2),
                     "Res-quad"   = c( 0, 0, 0, 0, 0,+2,-1,-2,-1,+2))


   
```


Run multcomp
```{r}
mult.comp.working <- glht(m.ANOVA.means       
                          ,linfct =contrst.mat
                          ,alternative = c("two.sided"))


mult.comp.summary <- summary(mult.comp.working
                             ,test = adjusted(type = "none"))
```


```{r}
print( summary(m.ANOVA.means),  correlation = FALSE)

```


### Save



```{r}
save.image(file = "./models/WRKSPC_age_rat_AOV_migVres_NOT_ind_spp.RData")
```
