---
title: "Clean raw Mencia pasture succession data"
author: "brouwern@gmail.com"
date: "February 16, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## General Workflow

Data was recieved from Steve Latta as seperate spreadsheets which were merge into a single .csv file in 0_proofing_merging....Rmd script

This script tidies/cleans up the data.

The following script makes final changes to prepare for analysis


## General Description of what this script does

See 0_proofing_merging....Rmd for more details

### SCRIPT 1: reshaping, tidying, and cleaning data
The following steps aren't necessarily all distinct in practice or necessarily occur in this precise order

#### Step 1.1: "reshaping" and "tidying"

##### "reshape"
taking data from the format it was entered in into the format that works best w/ R; usually reshaping from "wide" to "long" format.  eg, with data from each subplot in a different column to all data in a single column with subplot number indicated in a desperate column.

##### "tidying"
converting into an analysis-read format.  For example, splitting things like columns that combine treatment and plot number into separate columns.  (treatment column, subplot column)

#### Step 1.2: "cleaning"
Taking reshaped data and fixing any issues just as typos, incorrect column names, data entered into wrong field, fixing species names, etc.

After cleaning and tidying, the data could be "reshaped" (or recast) back into its original format and there would still be an ~1:1 correspondence with the datasheet.


## Libraries
```{r}
library(reshape2)
library(stringr)
library(stringi)
library(lubridate)
library(here)
```



## Load data


### Set up working directories 


### Load file

```{r}
filename <- here::here("data-raw",
                  "data-raw-mencia",
                  "mencia_site_persistence",
                  "CSV_mencia_site_persistence",
                 "mencia_all_b.csv")
print(filename)

skip <- 7
dat <- read.csv(file = filename,
                skip = skip,
                header =TRUE,
                strip.white = TRUE,
                blank.lines.skip = TRUE,
                na.strings = c("NA",""," "),
                stringsAsFactors = FALSE,
                colClasses=c("SUF"="character")
                )


```

### Check data

#### Check head and tail

Number blank columns to rigth and below spreadsheet
```{r}
head(dat); tail(dat)

```


#### Check summary
```{r}
summary(dat)

```



#### Check dim()

```{r}
dim(dat)

```

#### Change all headings to lower case

```{r}
names(dat) <- tolower(names(dat))
```



#### Reconcile names w/ other dataset

```{r}
names(dat[1:10])
names(dat) <- gsub("pre","band.pre",names(dat))
names(dat) <- gsub("suf","band.suf",names(dat))
```

#### Set up year variable

```{r}
summary(factor(dat$date))

dat$date2 <- dmy(dat$date)

dat$year <- year(dat$date2)
dat$month <- month(dat$date2)
dat$day <- day(dat$date2)
```


#### Move "died" from "wander" column to notes

```{r}
dat$notes <- ifelse(dat$wanderer == "DIED","DIED",NA)
```


#### Wander column

```{r}
summary(factor(dat$wanderer))
dat$wanderer <- gsub("X",NA,dat$wanderer)
dat$wanderer <- gsub("DIED",NA,dat$wanderer)
dat$wanderer <- factor(dat$wanderer)

summary(factor(dat$wanderer))
```

## Add "La" to sites

```{r}
dat$site <- gsub("Corral","El Corral",dat$site)
dat$site <- gsub("Caoba","La Caoba",dat$site)
dat$site <- gsub("Cueva","La Cueva",dat$site)
```



### Convert to factors

```{r}
dat$band.pre <- factor(dat$band.pre)
dat$band.suf <- factor(dat$band.suf)
dat$colors <- factor(dat$colors)
dat$age <- factor(dat$age)
dat$year.num <- dat$year
dat$year <- factor(dat$year)
dat$site <- factor(dat$site)
```



### Check out data
```{r}
summary(dat)
```

```{r}
summary(factor(dat$sex),1000)
```


```{r}
summary(factor(dat$notes),1000)
```


## Sites


### Fix typos

```{r}
dat$colors <- factor(gsub("x-","X-",as.character(dat$colors)))

dat$species <- factor(gsub("AM-KE","AMKE",as.character(dat$species)))

dat$sex <- factor(ifelse(dat$sex == "U", NA,dat$sex))
```





## Calculate age of sites


```{r}
site.age.tab <- data.frame(site = levels(dat$site),
                      site.age.init = c(20,5,2,10))

dat <- merge(dat,site.age.tab)
  
dat$site.age <- dat$site.age.init + 
                      dat$year.num -2003
```



## Create unique ID

```{r}
dat$band <- with(dat, paste(band.pre,
                                  band.suf,
                                  sep = "-"))
```















## Code spp

These are species judge to have sufficient sample sizes in 1st MS


### Focal migrants
```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

```

### Focal res
```{r}
focal.res <- c("HLCU"
,"STOF"
,"RLTH"
,"NOMO"
,"GRWA"
,"BANA"
,"BCPT"
,"YFGR"
,"BFGR"
,"GABU")




```

Code focal spp as resident or migrant
```{r}
i1 <- which(dat$species %in% focal.mig)
i2 <- which(dat$species %in% focal.res)

dat$status.focals <- NA
dat[i1, "status.focals"] <- "mig"
dat[i2, "status.focals"] <- "res"
```





### CHange "date" to "date.cap.orig"

```{r}
names(dat)[grep("date2",names(dat))] <- "date.cap.orig"
```




## Shorten names

```{r}
names(dat) <- gsub("resight","rs",names(dat))
names(dat) <- gsub("recap","rc",names(dat))
names(dat) <- gsub("20","",names(dat))
names(dat) <- gsub("nov","N",names(dat))
names(dat) <- gsub("jan","J",names(dat))
names(dat) <- gsub("feb","F",names(dat))
names(dat) <- gsub("wanderer","wander",names(dat))
names(dat) <- gsub("wander","wndr",names(dat))
names(dat) <- gsub("species","spp",names(dat))
names(dat) <- gsub("wing","wng",names(dat))
names(dat) <- gsub("tarsus","tar",names(dat))
names(dat) <- gsub("year","yr",names(dat))
names(dat) <- gsub("month","mo",names(dat))
names(dat) <- gsub("band","bnd",names(dat))
names(dat) <- gsub("sex","sx",names(dat))
names(dat) <- gsub("colors","col",names(dat))
names(dat) <- gsub("notes","nts",names(dat))
names(dat) <- gsub("status","stat",names(dat))
names(dat) <- gsub("age","ag",names(dat))
```

## Change "n/a" to "bf" for "before" 1st capture

Steve has "n/a" list for sampling occassions w/in a season before a bird was 1st observe (but not for years b/f)/

Change these to "bf"

```{r}
natobf<- function(x){ifelse(x == "n/a","bf",x)}
j.r <- grep("^r", names(dat))

dat[,j.r] <- apply(dat[,j.r],2, natobf)
```



## Change "X" to 1

steve list "X" for an observation or recapture.  
change to 1 to be consistent the mark recaput

```{r}
Xto1<- function(x){ifelse(x == "X" | x == "x","1",x)}
j.r <- grep("^r", names(dat))

dat[,j.r] <- apply(dat[,j.r],2, Xto1)
```

## Change site names to 1

steve list site names in some cases instead of "X".

Change to 1

```{r}
siteto1<- function(x){
  i <- which(x %in% c("Corral","Caoba","Cueva","Morelia","Cue/Cao",
                      "Fuera","DIED",
                      "repeated combo"))
  x[i] <- 1
  return(x)
  }

dat[,j.r] <- apply(dat[,j.r],2, siteto1)
```

## Convert all to factor
```{r}
for(i in 1:length(j.r)){
  j <- j.r[i]
  dat[,j] <- as.factor(dat[,j])
}
summary(dat[,j.r])
summary(factor(dat$rs.04.J))
```


## Calculate number of times captured

```{r}
calc_caps <- function(x){
  return(sum(as.numeric(x[grep("1",x)])))
}

dat$num.rc.rs <- apply(dat[,j.r],1, calc_caps)
```



```{r}
i.drop1 <- which(names(dat) %in% c("bnd.pre",#"bnd.suf",
                                       "wng","tar",
                                       "site.ag.init",
                                       "site.ag","bnd",
                                       "nts","yr.num",
                                       "date2","mo","day"))
                 
dat[,-c(j.r,i.drop1)]
```


## Were any wanders incorrectly coded?

None

```{r}
which(dat$wndr == "W" & dat$num.rc.rs > 0)
```

## Were any P (=persistent) never resighted or recaptured?

nope 

```{r}
which(dat$wndr == "P" & dat$num.rc.rs < 1)


```


## Are all residents coded 

Yes

```{r}
i.prob <- which(dat$wndr == "P" & dat$num.rc.rs < 1)

#dat[i.prob, -c(j.r,i.drop1)]
```




## Examine wanderer

```{r}
summary(factor(dat$wndr ))

i.w <- which(dat$wndr == "W")
j.r <- grep("^r", names(dat))

```


## Code "P"

```{r}
dat$persist <- ifelse(dat$num.rc.rs > 0, "P","NP")
```




## Change spp to spp.code

```{r}
names(dat) <- gsub("spp","spp.code",names(dat))
```


## Load spp meta / trait info

```{r}
spp.meta <- read.csv(here::here("data","spp_list.csv"),stringsAsFactors =F)
```


Subset columns I want to use
```{r}
spp.meta <- spp.meta[c("spp.code","spp",
                       "status2",
                       "hab1","diet")]

names(spp.meta) <- gsub("status2","stat",names(spp.meta))
```

```{r}
summary(spp.meta)
```

## Change GRWA -> GTGT

Green tailed ground warlber and green tailed ground tanager are the same

Change column name

```{r}
names(dat) <- gsub("species","spp.code", names(dat))
```


```{r}
dat$spp.code <- as.character(dat$spp.code)
dat$spp.code[which(dat$spp.code == "GRWA")] <- "GTGT"
```

```{r}
dat$spp.code[which(dat$spp.code == "HLCU")] <- "HILC"

```


MYWA =  myrtyle warbler = YRWA yellow rumped warbler
NUMA -> SBMU #nutmet manakin = scaly breasted munia
WHQD = WFQD

```{r}
dat$spp.code[which(dat$spp.code == "MYWA")] <- "YRWA"
dat$spp.code[which(dat$spp.code == "NUMA")] <- "SBMU"
dat$spp.code[which(dat$spp.code == "WHQD")] <- "WFQD"
```


## Merge cleaned data w/ species meta data


Merge
```{r}
dat2 <- merge(dat, spp.meta, all = T)

dim(dat)
dim(dat2)

dat2 <-dat2[-which(is.na(dat2$site) == TRUE), ]
dat <- dat2
```



## Save cleaned data in original form

### Save .csv
```{r}
filename <- "mencia_site_persist_clean.csv"
filename <- here("data-raw",
                  "data-raw-mencia",
                  "mencia_site_persistence",
                  filename)

write.csv(dat, file = filename,row.names = F)
```


```{r}

```

