---
title: "Analysis: sex & age ratio - regression"
author: "Nathan Brouwer"
date: "January 17, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Libraries
```{r}
library(lme4)
library(bbmle)
library(ggplot2)
library(cowplot)
```



## Load cleaned and scrubbed data


```{r}
load("./data/annual_captures.RData")

```


## Set up factors for easy interp with summary

```{r}

ann_caps$site <- factor(ann_caps$site)
ann_caps$spp.code <- factor(ann_caps$spp.code)
ann_caps$sex.M <- factor(ann_caps$sex.M,
                            levels = c("other","M"))
ann_caps$status.focals <- factor(ann_caps$status.focals)
summary(ann_caps[, c("spp.code","site",
                     "sex.M","age.AHY.01",
                     "site.age","site.age.init",
                     "status.focals")])


with(ann_caps, table(site,site.age,useNA = "always"))

with(ann_caps, table(age,age.AHY.01,useNA = "always"))
```


## Set arbitrary site age for Aceitillar

```{r}
ann_caps$site.age[which(ann_caps$site == "Aceitillar")] <- 50
```


## Subset by focal species

### Focal spp
```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

focal.res <- c("HLCU","STOF","RLTH"
,"NOMO","GRWA","BANA"
,"BCPT","YFGR","BFGR","GABU")

focals <- c(focal.mig,focal.res)
```


### Subset
```{r}
i.focals <- which(ann_caps$spp.code %in% focals)
 
ann_caps.focals <- ann_caps[i.focals,]
```


## Isolate unique individuals

To remove repeated measures; this simplifies the models but tosses out data

```{r}

i.unique <- match(unique(ann_caps.focals$band),
                  ann_caps.focals$band)

dim(ann_caps.focals)
length(unique(ann_caps.focals$band))
length(i.unique)


ann_caps.focals2 <- ann_caps.focals[i.unique,]
```




## Analysis of age ratio
Include ID?  most individuals only captured once; just use 1st time captures?


hab1
diet

```{r}
summary(factor(ann_caps.focals2$diet))
summary(factor(ann_caps.focals2$hab1))
```


```{r}

#model just those with real ages define?
i.use <- which(ann_caps.focals2$site != "Aceitillar")

summary(ann_caps.focals2$site)
summary(factor(ann_caps.focals2$year))
summary(factor(ann_caps.focals2$site.age))

m.age0 <- bglmer(age.AHY.01 ~ 1 +
               #(1|ID) +
               (1|year) +
               (1|site) +
               (1|spp.code),
              data = ann_caps.focals2[i.use,],
              family = binomial,fixef.prior = "t",
              glmerControl(optimizer = "Nelder_Mean"))



```


```{r}
#site age predictor
m.age1 <- update(m.age0, . ~ . + scale(site.age))

m.age1b <- update(m.age0, . ~ . + scale(site.age) + (scale(site.age)||spp.code))


m.age2 <- update(m.age0, . ~ . + hab1)

m.age3 <- update(m.age0, . ~ . + scale(site.age) + hab1)
m.age3b <- update(m.age0, . ~ . + scale(site.age) + hab1 + (site.age||spp.code))

m.age4  <- update(m.age0, . ~ . + scale(site.age)*hab1)
m.age4b <- update(m.age0, . ~ . + scale(site.age)*hab1 + (site.age||spp.code))
```

```{r}

```


```{r}
AICtab(m.age0,
       m.age1,# m.age1b,
       m.age2, 
       m.age3, #m.age3b,
       m.age4 #, #m.age4b
       )
```


```{r}
AICtab(m.age0,
       m.age1, m.age1b,
       m.age2, 
       m.age3, m.age3b,
       m.age4, m.age4b)
```


```{r}
plot_model(m.age4)
```


# plot Age Ratio

```{r}
pd <- position_dodge(0.1)
ggplot(data = ann_caps.focals[i.use,],
       aes(y = as.numeric(age.AHY.01)-1,
           x = site.age), 
       color = site,
       group = status.focals) +
  geom_jitter(aes(color = site), height = 0.1) +
  facet_wrap(~ status.focals) + 
  geom_smooth(method = glm,
              method.args = list(family = "binomial")) +
  ylab("Age Ratio") + theme_bw()

```




















## Analysis of sex ratio


```{r}
library(lme4)

ann_caps.focals2$ID <- with(ann_caps.focals2, paste(band, spp.code))

#i.use <- which(ann_caps.focals2$site.age > 0)



dim(ann_caps.focals)
length(i.use)
m0 <- glmer(sex.M ~ 1 +
               #(1|ID) +
               (1|year) +
               (1|site) +
               (1|spp.code),
              data = ann_caps.focals2[i.use,],
              family = binomial)

```


```{r}
m1 <- update(m0, . ~ . + site.age)
m2 <- update(m0, . ~ . +            status.focals)
m3 <- update(m0, . ~ . + site.age + status.focals)
m4 <- update(m0, . ~ . +                          site.age*status.focals)
```


```{r}
AICtab(m0, m1, m2, m3, m4)
```

```{r}
summary(m3)
```

