---
title: "Community Level analyses: Species Richness"
author: "Nathan Brouwer"
date: "February 9, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Function that calls all libraries needed for analysis
```{r}
source("./R/load_libraries.R")
load_libraries()
```


## Load data

Dataframe contains the following variables for each site and year

* simpson's index
* spp.rich
* evenness

"dominance" also reported in the original MS but not calculate as of 2/12/18

```{r}
load(file = "./data/community_dat.RData")
```


## Plot data

```{r}
# library(ggpubr)
# 
# ggplot(data = community_dat,
#             aes(y = spp.rich,
#             x = site.age)) +
#   geom_point() +
#   geom_smooth(method = "lm")
```

## Model

### Model Pasture sites

#### Log transform for inference
```{r}
i.aceit <- which(community_dat$site == "Aceitillar")
#community_dat$i <- 1:dim(community_dat)[1]

#null model
m.even.0.logit <- blmer(logit(evenness) ~ 1 + 
                       #(1|i) +
                       (1|site) +
                       (1|year),
                     #family = "poisson",
                     data = community_dat[-i.aceit,])

#site model
m.even.site.logit  <- update(m.even.0.logit, . ~ . + site.age.cent)

#results
anova(m.even.0.logit,
      m.even.site.logit)

summary(m.even.site.logit)
```


#### un transform for plotting inference
```{r}
i.aceit <- which(community_dat$site == "Aceitillar")
#community_dat$i <- 1:dim(community_dat)[1]

#null model
m.even.0 <- blmer(evenness ~ 1 + 
                       #(1|i) +
                       (1|site) +
                       (1|year),
                     #family = "poisson",
                     data = community_dat[-i.aceit,])

#site model
m.even.site <- update(m.even.0, . ~ . + site.age.cent)

#results
anova(m.even.0,
      m.even.site)

summary(m.even.site)
```


### Model Aceitillar alone



### 1-way ANOVA style model

#### Remove overlap

```{r}
i.caoba.drop.comm <- which(community_dat$site == "La Caoba" & 
                      community_dat$site.age < 7) #keep age 7, 8, 9
i.cueva.drop.comm <- which(community_dat$site == "La Cueva" & 
                      community_dat$site.age > 4) #keep age 2, 3, 4


community_dat_2 <- community_dat[-c(i.cueva.drop.comm,
                                    i.caoba.drop.comm), ]
```

#### Run models - original scale
```{r}

#null - w/ ranef for site
m.anova.even.0 <- blmer(evenness ~ 1 + 
                       #(1|i) +
                       (1|site) +
                       (1|year),
                     #family = "poisson",
                     data = community_dat_2[,])

# site
m.anova.even.site <- blmer(evenness ~ site + 
                       #(1|i) +
                       #(1|site) +
                       (1|year),
                     #family = "poisson",
                     data = community_dat_2[,])
# means
m.anova.even.means <- blmer(evenness ~ -1 + site + 
                       #(1|i) +
                       #(1|site) +
                       (1|year),
                    # family = "poisson",
                     data = community_dat_2[,])


```

#### Inference

##### Model comparisons

```{r}
ICtab(m.anova.even.0,
      m.anova.even.site,
      type = "AICc",
      weights = T,
      delta = T)
```


##### Contrast modeling

Create contrast matrix
```{r}
contrst.mat.comm <- rbind("Linear" = c(-2,-1, 0, 1, 2),
                          "Quad"   = c(+2,-1,-2,-1,+2))
```


Run multcomp
```{r}
m.anova.even.means.logit <- blmer(logit(evenness) ~ -1 + site + 
                       #(1|i) +
                       #(1|site) +
                       (1|year),
                    # family = "poisson",
                     data = community_dat_2[,])
mult.comp.working.even <- glht(m.anova.even.means.logit       
                          ,linfct =contrst.mat.comm
                          ,alternative = c("two.sided"))


mult.comp.summary.even <- summary(mult.comp.working.even
                             ,test = adjusted(type = "none"))
```

Mult comp output

```{r}
mult.comp.summary.even
```





## Plotting Simpsons's diversity

### Plot pasture sites

#### Calculate CIs for pasture sites

```{r}
#merTools::predictInterval()
source('~/1_R/git/git-aviary/DRmencia/R/build_CIs_regrxn_pois_comm.R')
even.rgrxn.preds <- buidCIs_regrxn_pois_comm(m.even.site,mod.type = "normal")

```


Determine mean of site.age variable to de-center it while plotting
```{r}
covar.mean <- mean(community_dat$site.age.init,
                   na.rm = TRUE)
```

#### Calculate CIs for ANOVA
```{r}
source('~/1_R/git/git-aviary/DRmencia/R/buidCIs_ANOVA_pois_comm.R')
even.ANOVA.preds <- buidCIs_ANOVA_pois_comm(mod = m.anova.even.site,
                                           mod.type = "normal")

```

#### Build regression Plot
```{r}
source('~/1_R/git/git-aviary/DRmencia/R/back_transform_reals.R')
source('~/1_R/git/git-aviary/DRmencia/R/plot_regrxn_pois_comm.R')

even.rgrxn.preds <- back_transform_reals(even.rgrxn.preds,fnxn = "none")
even.ANOVA.preds <- back_transform_reals(even.ANOVA.preds,fnxn = "none")


regrxnplot_even <- plot_regrxn_pois_comm(dat = even.rgrxn.preds,
                      ANOVA.dat = even.ANOVA.preds, 
                      mod = m.even.site,
                      mean.of.covar = covar.mean,
                      ylab. = "Eveness",
                      ymin. = 0.6,
                      ymax = 1)
```

#### Build Aceitillar plot
```{r}
i.aceit <- which(even.ANOVA.preds$site == "Aceitillar")
i.aceit.mod <- which(m.anova.even.site@frame$site == "Aceitillar")
even.ANOVA.preds <- back_transform_reals(even.ANOVA.preds,fnxn = "none")

source('~/1_R/git/git-aviary/DRmencia/R/plot_ANOVA_pois_comm.R')
anovaplot_even <- plot_ANOVA_pois_comm(even.ANOVA.preds[i.aceit,],
                                               mod = m.anova.even.site, 
                                               mod.dat.i = i.aceit.mod,
                                               ylab. = "",
                                               point.sz = 4,
                                           ymin. = 0.6,
                                           ymax = 1)


```


## Save constituent plots
```{r}
save.image("./plots/regrxnplot_even.RData")
save.image("./plots/anovaplot_even.RData")
```



## Plot both Regression and ANOVA together

```{r}
cowplot::plot_grid(regrxnplot_even,
                   anovaplot_even,
                   rel_widths = c(8,1))
```


```{r}

```










