---
title: "Body condition analysis"
author: "Nathan Brouwer"
date: "February 5, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Species w/ sig results in original MS
a)	Bananaquit  
b)	Black-crowned Palm-Tanager  
c)	Greater Antillean Bullfinch (male)
d)	Greater Antillean Bullfinch (female)
e)	Green-tailed Ground-Tanager


## Libraries

```{r}
library(here)
fi <- here::here("R","load_libraries.R")
source(fi)
load_libraries()
```

## Load data

```{r}
load(here::here("data","condition.RData"))
```


## Prep data
```{r}
condition$wing.log <- log(condition$wing)
condition$mass.log <- log(condition$mass)
```

##

```{r}
with(condition, table(spp.code, 
                      site))
```


## Plot raw data

```{r}
condition$group <- with(condition,
                      paste(spp.code, site))

#outlier in original analysis
i.use <- which(is.na(condition$stat.focals) == FALSE)
length(i.use)
condition$outlier <- "ok"
condition$outlier[c(551, 731, 2404, 1608)] <- "outlier"
i.use <- which(is.na(condition$stat.focals) == FALSE)
condition.plot <- condition[i.use,]

i.mig <- which(condition.plot$stat.focals == "mig")
i.res.in.aceit <- which(condition.plot$spp.code %in% c("BANA","BCPT","GABU","GRWA","STOF"))

with(condition.plot[-i.mig, ],
     table(spp.code, site.age))

ggplot(data = condition.plot[i.res.in.aceit, ],
       aes(y = log(mass),
           x = log(wing),
           #color = spp.code,
           group = group)) +
   facet_wrap(~ spp.code,
              scales = "free") +
  geom_point(aes(color = site.age,
                  shape = site.age))  +
  geom_smooth(method = lm,
              formula = y ~ x , 
              se = FALSE,
              aes(color = site.age))# +
   #geom_smooth(se = FALSE) +
    
```

## Examine focal  species

```{r}
i.foc <- i.BANA <- which(condition$spp.code == "BANA")

```


### ANCOVA plot for focal spp
```{r}

with(condition.plot[i.foc, ],
     table(site))
ggplot(data = condition.plot[i.foc, ],
       aes(y = log(mass),
           x = log(wing),
           color = site.age,
           group = site.age)) +
  geom_point(aes(color = site.age,
                  shape = site.age))  +
  geom_smooth(method = lm,
              formula = y ~ x , 
              se = T,
              aes(color = site.age))# +
```


### Plot ratio
Radios used in initial analysis by SL
```{r}
condition.plot$site.age <- factor(condition.plot$site.age,
                                  levels = c("2","5","10","20",
                                             "mature"))
ggplot(data = condition.plot[i.foc, ],
       aes(y = mass/wing,
           x = site.age,
           color = site.age,
           group = site.age)) +
  # geom_point(aes(color = site.age,
  #                 shape = site.age))   +
 # geom_boxplot
  stat_summary(fun.data = "mean_cl_boot",
               colour = "red", size = 2)


with(condition.plot[i.foc, ],
     hist(mass/wing))
```



### Model ratio

```{r}
m.ratio.0 <- lm(mass/wing ~ 1, data = , condition.plot[i.foc, ])
m.ratio.site <- update(m.ratio.0, . ~ . + site.age)

anova(m.ratio.0,m.ratio.site)

aov.ratio.site <- aov(mass/wing ~ site.age, data= condition.plot[i.foc, ])

TukeyHSD(aov.ratio.site)
```


### Model ANCOVA

```{r}
m.ancova.0 <- lm(mass ~ wing, data = condition.plot[i.foc, ])
m.ancova.site <- update(m.ancova.0, . ~ . + site.age)

anova(m.ancova.0,m.ancova.site)
```


### Model ANCOVA w/ranefs

```{r}
m.lmer.0 <- lmer(mass ~ wing + (1|site), data = condition.plot[i.foc, ])
m.lmer.site <- update(m.lmer.0, . ~ . + site.age)

anova(m.lmer.0,m.lmer.site)
```





## Use condition indicies

## RMA

```{r}
#download.packages("smatr")
library(smatr)
```

```{r}
i.BANA <- which(condition$spp.code == "BANA")
i.BCPT <- which(condition$spp.code == "BCPT")
i.GRWA <- which(condition$spp.code == "GRWA")
```


## Run sma

Major axis regression
```{r}
i.wrkng <- i.GRWA
sma.1 <- sma(mass ~ wing , data = condition.plot[i.wrkng, ],
    log="xy", method=c("SMA"), 
    type=c("elevation"),
    multcomp=FALSE, multcompmethod=c("default","adjusted"),
    robust=FALSE)
```

Calculate mean size variable
```{r}
L0 <- mean(log(condition.plot[i.wrkng, "wing"]))
```

Function to calculate M.hat
```{r}
M.hat <- function(Mi, Li, L0,b.SMA){
  Mi*((L0/Li)^b.SMA)
}
```

Calcualte M.hat values
```{r}
M.hat.i <- M.hat(Mi = condition.plot[i.wrkng, "mass"],
      Li = log(condition.plot[i.wrkng, "wing"]),
      L0 = log(mean(condition.plot[i.wrkng, "wing"])),
      b.SMA = coef(sma.1)[["slope"]])
```



Build dataframe
```{r}
#combine orig data with M.hat
M.hat.out <- cbind(condition.plot[i.wrkng, ],
      M.hat.i)

#Calculat residuals
M.hat.out$ei <- resid(lm(log(mass) ~ log(wing), data = condition.plot[i.wrkng, ]))
```

```{r}
pt.sz <- 1
#Plot M hat
p1 <- ggplot(data = M.hat.out,
       aes(y = M.hat.i,
           x = site.age)) +
  stat_summary(fun.data = "mean_cl_boot",
               colour = "red", size =pt.sz) +
  ggtitle("Scaled mass index")

#Plot ratios
p2 <- ggplot(data = M.hat.out,
       aes(y = mass/wing,
           x = site.age)) +
  stat_summary(fun.data = "mean_cl_boot",
               colour = "red", size = pt.sz)+
  ggtitle("ratio mass:wing")

#Plot residuals
p3 <- ggplot(data = M.hat.out,
       aes(y = ei,
           x = site.age)) +
  stat_summary(fun.data = "mean_cl_boot",
               colour = "red", size = pt.sz)+
  ggtitle("residuals mass~wing")

cowplot::plot_grid(p1, p2,
                   p3)
```


















