---
title: "Finalize 2016 herb data for analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Reshape to wide, calculate summaries, etc


## Libraries
```{r}
library(here)
library(reshape2)
library(car)
library(stringi)
```



## Load cleaned and scrubbed data

Data has been cleaned to removed typos and basic prep done for statistics

```{r}
filename <- "mencia_cleaned.csv"
filename <- here("data-raw",
                  "data-raw-mencia",
                  "mencia_all_captures_by_year",
                  filename)

mencia <- read.csv(filename)
```


## Order sites by age

```{r}
mencia$site <- factor(mencia$site,
                      levels = c("La Cueva",
                                 "La Caoba",
                                 "Morelia",
                                 "El Corral"))
```



## Cast for demographic summaries

Identify first instance of bird occuring.  (I checked - all birds have the same age code even if recaptured; recaps rare regardless)
```{r}
i.unique <- match(unique(mencia$band),mencia$band)

length(which(duplicated(mencia$band) == TRUE))
```


### cast by age

```{r}
age.cast <- dcast(data= mencia[i.unique,],
      formula = species + status ~ site,
      value.var = "site",
      fun.aggregate = length)


age.cast[which(age.cast$status == "mig"), ]
```





### cast by sex

```{r}
sex.cast <- dcast(data= mencia[i.unique,],
      formula = species + status  ~ site + sex,
      value.var = "site",
      fun.aggregate = length)


sex.cast[which(sex.cast$status == "mig"), ]
```







## Save output

### save .csv files

save to raw data folder
```{r}
# focal spp and summary output
# filename <- "focals_herbdat_16.csv"
# filename <-here("data-raw","data-raw-herb","herb_veg_16",
#       filename)
# write.csv(focals.herbdat.16, 
#           file = filename,row.names = F)
# 
# 
# #cover by species
# filename <- "cover_by_spp_herbdat_16.csv"
# filename <-here("data-raw","data-raw-herb","herb_veg_16",
#       filename)
# write.csv(cover.by.spp.herbdat.16, 
#           file = filename,row.names = F)
# 
# 
# 
# #pres abs by spp
# filename <- "presabs_by_spp_herbdat_16.csv"
# filename <-here("data-raw","data-raw-herb","herb_veg_16",
#       filename)
# write.csv(presabs.by.spp.herbdat.16, 
#           file = filename,row.names = F)


```


### save .RData files

saved to "data" folder where working data for real analyses is stored

```{r}
# filename <- "focals_herbdat_16.RData"
# filename <-here("data",
#       filename)
# save(focals.herbdat.16, 
#           file = filename,row.names = F)
```



```{r}

```


