---
title: "Finalize 2016 herb data for analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Do final creation of synthetic varibles

## Libraries
```{r}
library(here)
library(reshape2)
library(car)
library(stringi)
library(tibble)
```



## Load cleaned and scrubbed data

Data has been cleaned to removed typos and basic prep done for statistics

```{r}
filename <- "mencia_cleaned.csv"
filename <- here::here("data-raw",
                  "data-raw-mencia",
                  "mencia_all_captures_by_year",
                  filename)

mencia <- read.csv(filename)
```


## Order sites by age

```{r}
mencia$site <- factor(mencia$site,
                      levels = c("La Cueva",
                                 "La Caoba",
                                 "Morelia",
                                 "El Corral"))
```




## Make unique annual ID

ID is unique for each individual
some individuals captured multiple times w/in years
some re-captured between years.
Make unique ID to seperate out w/in year recaps

```{r}
mencia$ID.yr <- with(mencia,
                     paste(band,
                           year),
                     sep = ".")


```


447 re-captured in different sites

```{r}
length(unique(mencia$band))  #3143
length(unique(mencia$ID.yr)) #3590
length(unique(mencia$ID.yr)) - length(unique(mencia$band))
```

### Make unique site ID

161 re-captured in different sites (w/in yr OR between yr)
```{r}
mencia$ID.site <- with(mencia,
                     paste(band,
                           site),
                     sep = ".")

length(unique(mencia$ID.site))-length(unique(mencia$band))
```


### Make Male/not male code

```{r}
mencia$sex.M <- with(mencia,
                     ifelse(sex == "M","M","other"))
# has to be sexed to count
#mencia$sex.M[is.na(mencia$sex.M)] <- "other"
mencia$sex.M <- factor(mencia$sex.M)
```


### Make AHY/not AHY code


Combine all AHY and beyond catetories

Has to be aged to count

```{r}
summary(mencia$age)
mencia$age <- as.character(mencia$age)

#HY and SY = "age1"
mencia$age.AHY.01 <- gsub("^HY","age1", mencia$age)
mencia$age.AHY.01 <- gsub("^SY","age1",mencia$age.AHY.01)

mencia$age.AHY.01 <- gsub("^AHY","age2",mencia$age.AHY.01)
mencia$age.AHY.01 <- gsub("^ASY","age2",mencia$age.AHY.01)
   

mencia$age.AHY.01  <- factor(mencia$age.AHY.01 )

summary(mencia$age.AHY.01 )
```


## Save finalized data

### Save .csv
```{r}
filename <- "mencia_annual_final.csv"
filename <- here::here("data-raw",
                  filename)

write.csv(mencia, file = filename,row.names = F)
```



```{r}

```


