---
title: "Analysis: Bird Abundance - Poisson GLMM regression - Pasture sites"
author: "Nathan Brouwer"
date: "January 17, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The script carries out an analysis of the abundance trends of each species using a regression style GLMM and extracting random slopes for each species.  It loads a dataframe with the following columns

"year.num"      "site"          "spp.code"      "hab1"          "diet"         
"year"          "site.age.cent" "status.focals" "N"             "i"            
"site.age.init" "site.age"    


## Preliminaries

### Load cleaned, merged and scrubbed data

```{r}
load(file = "./data/ann_counts.RData") #loads "ann_counts"
```


### Check for any bad entries
```{r}
levels(ann_counts$site) 

summary(net_hrs$year)
summary(factor(ann_counts$year))
intersect(names(net_hrs),
          names(ann_counts))
```



## Load libraries

Function that calls all libraries needed for analysis
```{r}
source("./R/load_libraries.R")
load_libraries()
```

## Look at data

```{r}
names(ann_counts)
```


## Index for Removing aceitillar

Aceitillar is run seperately using 1-way ANOVA style models in the next script in this series.

```{r}
i.N.aceit <- which(ann_counts$site == "Aceitillar")
```


## Model counts w/Poisson GLMM

### Model random slopes

* Full random slopes model via (site.age.cent|spp.code).  
* Note (1|i) is a observation-level random effect to accomodate overdispersion.
* Aceitillar dropped
* Runs w/o convergence warning

```{r m.N.rand.slopes}
m.N.rand.slopes <- bglmer(N ~ 1 +
               (1|i) + 
               (1|year) +
               (1|site) +
               (site.age.cent|spp.code),
               offset = log(net.hours),
              data = ann_counts[-i.N.aceit, ],
              family = poisson ,
              glmerControl(optimizer = "Nelder_Mead")
              )
```




### Extract blups and calculate SEs

* BLUPS extracted with arm::se.coef

```{r}
pois.blup.out <- data.frame(beta = ranef(m.N.rand.slopes)[["spp.code"]]$site.age.cent,
se  = se.coef(m.N.rand.slopes)["spp.code"][["spp.code"]][,"site.age.cent"])

pois.blup.out$Species <- row.names(pois.blup.out) 
```


### Plot random slopes

Set up names for labeling
```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW",
               "MAWA","NOPA","PALM") #add to poisson

focal.res <- c("HLCU","HILC",         #fix - which name to use?
               "STOF",
               "RLTH","NOMO",
               "GRWA",       "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "GABU",
               "BWVI", #32; Aceitillar / Cueva +misc
               "GREP") #18 just Aceitillar - ever yr

focals <- c(focal.mig,focal.res)
```


Label residents vs. migrants
```{r}
pois.blup.out$Species <- factor(pois.blup.out$Species)
pois.blup.out$Status <- "Resident"
pois.blup.out$Status[which(pois.blup.out$Species %in% focal.mig)] <- "Migrant"
```


## Plot trends

To Do:
-add endemic status
-plot against observed abundance
-NOTE: several spp. have very non-linear pattern!
-sort by magnitude
-add info on CIs

```{r}
ggplot(dat = pois.blup.out,
       aes(y = exp(beta),
           x = Species,
           color = Status)) +
  geom_errorbar(aes(ymax = exp(beta+1*se),
                    ymin = exp(beta-1*se)),
                width = 0,
                size =2) +
  geom_errorbar(aes(ymax = exp(beta+1.96*se),
                    ymin = exp(beta-1.96*se)),
                width = 0) +
      geom_point(size = 4) +
  geom_hline(yintercept = exp(0)) +
  coord_flip() +
  facet_wrap(~Status,scales = "free") +
  ylab("Trend") 
```


```{r}

```









```{r}

```





