---
title: "Analysis: Bird Abundance - Poisson GLMM 1-way ANOVA - All sites"
author: "Nathan Brouwer"
date: "February 13, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The script carries out an analysis of the abundance trends of each species using a 1-way ANOVA style GLMM using ANOVA linear and quadratic trend tests. 

The script loads a dataframe with the following columns (same as for regression style GLMM)

"year.num"      "site"          "spp.code"      "hab1"          "diet"         
"year"          "site.age.cent" "status.focals" "N"             "i"            
"site.age.init" "site.age"    


### Preliminary results
Text related to these results from email to Steve latta
patterns "often complex.  BAWW shows a consistent increase from the pasture sites continuing on to Aceitillar.  COYE are PAWA are both abundant at a single pasture site.  OVEN and BTBW both show overall linear trends when you just look at the pasture sites, but if you consider Aceitillar there is evidence that their abundance actually peaks for the middle-aged pasture.]"

### Load cleaned, merged and scrubbed data

 
```{r}
load(file = "./data/ann_counts.RData") #loads "ann_counts"
```


#### Remove overlap

Caoba and Cueva both occur at same ages; overlap removed b/c "site" is being treated as a categorical variable in a 1-way ANOVA.

```{r}
#younger site
i.cueva.drop.N <- which(ann_counts$site == "La Cueva" & 
                      ann_counts$site.age > 4) #keep age 2, 3, 4

#older site
i.caoba.drop.N <- which(ann_counts$site == "La Caoba" & 
                      ann_counts$site.age < 7) #keep age 7, 8, 9

#subset data
ann_counts_2 <- ann_counts[-c(i.cueva.drop.N,
                                    i.caoba.drop.N), ]
```




## Contrast modeling

### Example run: single species

* means coding for each level
* (1|i) is observation-level random effect for overdispersion

```{r}
i.OVEN <- which(ann_counts_2$spp.code == "OVEN")
i.AMRE <- which(ann_counts_2$spp.code == "AMRE")

summary(ann_counts_2[i.OVEN, ])

library(doBy)
summaryBy(N/net.hours*1000 ~ site+year, data = ann_counts_2[i.AMRE, ],FUN = c(mean,sd))
summaryBy(N/net.hours*1000 ~ site+year, data = ann_counts_2[i.AMRE, ],FUN = c(sd),na.rm)
summaryBy(N/net.hours*1000 ~ site+year, data = ann_counts_2[i.OVEN, ],FUN = c(mean,sd))

head(ann_counts_2[i.OVEN, ])
with(ann_counts_2[i.OVEN, ], table(year,site))
m.anova.test <- bglmer(N ~ -1 + site +
               (1|i) + 
               (1|year),
               offset = log(net.hours),
              data = ann_counts_2[i.OVEN, ],
              family = poisson ,
              glmerControl(optimizer = "Nelder_Mead")
              )

summary(m.anova.test)
```


#### Create contrast matrix for trend analyses
```{r}
contrst.mat.comm.5 <- rbind("Linear" = c(-2,-1, 0, 1, 2),
                          "Quad"   = c(+2,-1,-2,-1,+2))

contrst.mat.comm.4 <- rbind("Linear1" = c(-3,-1, 1, 3),
                          "Quad"   = c(1,-1,-1,1))
```

#### Run multcomp
```{r}
# run multcomp::glht to get contrats
mult.comp.OVEN <- glht(m.anova.OVEN       
                          ,linfct =contrst.mat.comm.5
                          ,alternative = c("two.sided"))

# extract output
mult.comp.summary.OVEN <- summary(mult.comp.OVEN
                             ,test = adjusted(type = "none"))
```


#### Mult comp output

```{r}
mult.comp.summary.OVEN
```



### Contrast modeling: All species

#### Focal species

Species with sparcer counts are more difficult to model b/c they may not occur at all sites or may be very rare in some years.  

```{r}
focal.mig.anova <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW"#,
               #"MAWA",
               #"NOPA",
               #"PALM"
               ) #add to poisson

focal.res.anova <- c( #"HLCU","HILC",         #fix - which name to use?
                #"STOF",
                "RLTH",
               "NOMO",
                #"GRWA",     
                "HILC",
                "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "STOF",
               "GABU",
               "BWVI"#, #32; Aceitillar / Cueva +misc
               #"GREP"
               ) #18 just Aceitillar - ever yr

focals.anova <- c(focal.mig.anova,focal.res.anova)
```



### Examine species to be used in models
```{r}
#load raw annual capture data
load("./data/ann_caps.RData")

tot.obs.by.hab <- dcast(data = ann_caps,
      formula = spp.code  ~  site,
      value.var = "spp.code",
      fun.aggregate = length)


i. <- which(tot.obs.by.hab$spp.code %in% focals.anova)
tot.obs.by.hab[i. ,]


```



### Loop over each focal spp
```{r}
#focal species to use
spp.to.use.anova <- focal.mig.anova

#create list
multcomp.list <- as.list(spp.to.use.anova)
names(multcomp.list) <- spp.to.use.anova
options(warn = -1)

for(i in 1:length(spp.to.use.anova)){
  
i.spp <- which(ann_counts_2$spp.code == spp.to.use.anova[i])

m.anova.spp <- bglmer(N ~ -1 + site +
               (1|i) + 
               (1|year),
               offset = log(net.hours),
              data = ann_counts_2[i.spp, ],
              family = poisson ,
              glmerControl(optimizer = "Nelder_Mead")
              )

#some spp did not occur at every site and so
##cannot be evaluated
print(i)
print(spp.to.use.anova[i])

N.sites <- length(fixef(m.anova.spp) == 5)                          
summary(m.anova.spp)
print(N.sites)

if(N.sites == 5) {
mult.comp.spp <- glht(m.anova.spp       
                          ,linfct =contrst.mat.comm.5
                          ,alternative = c("two.sided"))
} 

if(N.sites == 4){
  mult.comp.spp <- glht(m.anova.spp       
                          ,linfct =contrst.mat.comm.4
                          ,alternative = c("two.sided"))
} 


mult.comp.summary.spp <- summary(mult.comp.spp
                             ,test = adjusted(type = "none")) 

print(mult.comp.summary.spp)

multcomp.list[[i]] <- mult.comp.summary.spp

}


str(multcomp.list,1)

```



#### Function to extract info from multcomp object

```{r}
source('~/1_R/git/git-aviary/DRmencia/R/getmultcompstats.R')

```


```{r}

## original function

      # mc.out<- multcomp.list[[3]]
      # 
      # getmultcompstats <- function(mc.out,
      #                              dist = "poisson"){
      #   
      #   #extract each part of output for a single model
      #   cs<-mc.out$test$coefficients
      #   ts<-mc.out$test$tstat
      #   ps<-mc.out$test$pvalues
      #   
      #   #put into vector
      #   x <- c(cs,ts,ps)
      #   
      #   #name vector components
      #   names(x) <- paste( names(x),
      #                      c("B","B","t","t","p","p"),sep = ".")
      #   
      #   #exp poisson data
      #   if(dist == "poisson"){
      #      mc.means <- exp(mc.out$coef)
      #   }
      #   
      #   if(dist != "poisson"){
      #      mc.means <- mc.out$coef
      #   }
      #  
      #  
      #   #vector of ALL site names
      #   all.sites <- c("siteLa Cueva","siteLa Caoba",
      #                  "siteMorelia","siteEl Corral",
      #                  "siteAceitillar" )
      #   
      #   #blank vector
      #   y <- rep(NA,5)
      #   names(y) <- all.sites
      #   
      #   #match up output (has to be done b/c sometimes there are 4 sites
      #   ## and sometimes there are 5 sites)
      #   ## very clunk but gets it done
      #   for(i in 1:length(mc.means)){
      #     j <- which(names(mc.means)[i] == names(y))
      #     y[j] <-  mc.means[i]
      #   }
      #   
      #   
      #   return(c(x,y))
      # }


```

Extract results for one line
```{r}
#debugonce(getmultcompstats)
getmultcompstats(multcomp.list[[1]])


```


Data frame to hold all results
```{r}
mcomp.stats.tab <- data.frame(Species = names(multcomp.list),
          #model output                              
           Linear.B = NA,
           Quad.B = NA,
           Linear.t = NA,
           Quad.t = NA,
           Linear.p = NA,
           Quad.p = NA,
           Linear.SE = NA,
           Quad.SE = NA,
          
          #site means
           siteLaCueva=NA,
           siteLaCaoba =NA,
           siteMorelia  =NA,
           siteEl.Corral=NA,
           siteAceitillar = NA,
          
          #CI lo
           CI.lo.LaCueva=NA,
           CI.lo.LaCaoba =NA,
           CI.lo.Morelia  =NA,
           CI.lo.El.Corral=NA,
           CI.lo.Aceitillar = NA,
          
           #CI hi
           CI.hi.LaCueva=NA,
           CI.hi.LaCaoba =NA,
           CI.hi.Morelia  =NA,
           CI.hi.El.Corral=NA,
           CI.hi.Aceitillar = NA)

```


### Loop over and extract
```{r}
for(i in 1:length(multcomp.list)){
  print(i)
  mcomp.stats.tab[i,-1] <- getmultcompstats(multcomp.list[[i]])
}

```
 

### Set order
```{r}
library(wrapr)
j.names <- qc(Species,
              Linear.B,Linear.SE,Linear.t,Linear.p,
   Quad.B,  Quad.SE,  Quad.t,  Quad.p,
   siteLaCueva, CI.lo.LaCueva, CI.hi.LaCueva,
   siteLaCaoba, CI.lo.LaCaoba, CI.hi.LaCaoba,
   siteMorelia, CI.lo.Morelia, CI.hi.Morelia,
   siteEl.Corral, CI.lo.El.Corral, CI.hi.El.Corral,
   siteAceitillar, CI.lo.Aceitillar, CI.hi.Aceitillar)
mcomp.stats.tab2 <- mcomp.stats.tab[,j.names]


names(mcomp.stats.tab2) <- gsub("site","",names(mcomp.stats.tab2))
names(mcomp.stats.tab2) <- gsub("^La","La ",names(mcomp.stats.tab2))
names(mcomp.stats.tab2) <- gsub("^El.","El ",names(mcomp.stats.tab2))
```


## Save table
```{r}
mcomp.stats.tab.round <- mcomp.stats.tab2
mcomp.stats.tab.round[,-1] <- round(mcomp.stats.tab.round[,-1],3)
write.csv(mcomp.stats.tab.round,file = "./tables/tables_total_captures_spp_specific_ANOVA_trends/ind_spp_trends_with_Aceitillar_update.csv",row.names = F)
```


## Make plot

```{r}
j.names <- qc(
   LaCueva, CI.lo.LaCueva, CI.hi.LaCueva,
   LaCaoba, CI.lo.LaCaoba, CI.hi.LaCaoba,
   Morelia, CI.lo.Morelia, CI.hi.Morelia,
   ElCorral, CI.lo.El.Corral, CI.hi.El.Corral,
   Aceitillar, CI.lo.Aceitillar, CI.hi.Aceitillar)
j.names <- gsub("^La", "La ",j.names)
j.names <- gsub("^El", "El ",j.names)

mcomp.melt <- melt(data = mcomp.stats.tab2,
                   id.vars = "Species",
                   measure.vars = j.names)

#create site variable
mcomp.melt$Site <- mcomp.melt$variable
mcomp.melt$Site <- gsub("CI.lo.","",mcomp.melt$Site)
mcomp.melt$Site <- gsub("CI.hi.","",mcomp.melt$Site)
mcomp.melt$Site <- gsub("^La", "La ",mcomp.melt$Site)
mcomp.melt$Site <- gsub("^El", "El ",mcomp.melt$Site)
mcomp.melt$Site <- gsub("^El .Corral", "El Corral ",mcomp.melt$Site)

mcomp.melt$Site <- gsub("  ", " ",mcomp.melt$Site)


#create type of var column
mcomp.melt$variable2 <- mcomp.melt$variable
mcomp.melt$variable2 <- gsub("(CI.lo.)(.*)","\\1",mcomp.melt$variable2)
mcomp.melt$variable2 <- gsub("(CI.hi.)(.*)","\\1",mcomp.melt$variable2)
mcomp.melt$variable2[-grep("CI",mcomp.melt$variable2)] <- "mean"
```

Reshape
```{r}

tab.forplotting <- dcast(data = mcomp.melt,
      formula = Species + Site ~ variable2,
      value.var = "value",
      fun.aggregate = mean)

mcomp.melt[which(mcomp.melt$Species == "PAWA"), ]
tab.forplotting[which(tab.forplotting$Species == "PAWA"), ]
```



## Plot

```{r}
tab.forplotting$Site <- gsub(" $","",tab.forplotting$Site)
tab.forplotting$Site2 <- gsub("La ","",tab.forplotting$Site)
tab.forplotting$Site2 <- gsub("El ","",tab.forplotting$Site2)
summary(factor(tab.forplotting$Site2))

tab.forplotting$Site2 <- factor(tab.forplotting$Site2,
        levels = c("Cueva",
                   "Caoba",
                   "Morelia",
                   "Corral",
                   "Aceitillar"))

ggplot(data = tab.forplotting,
       aes(y = mean*1000,
           x = Site2,
           group = Species)) +
  geom_point() +
  geom_line() +
  xlab("Site") +
  ylab("Captures per 1000 net hours") +
  geom_errorbar(aes(ymax = CI.hi.*1000,
                    ymin = CI.lo.*1000),
                width = 0) +
  facet_wrap(~Species,
             scale = "free")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r}

```

