---
title: "Analysis: Bird Abundance - Poisson GLMM with quadratic terms - pasture sites"
author: "Nathan Brouwer"
date: "February 21, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

Script loads a dataframe with the following columns

"year.num"      "site"          "spp.code"      "hab1"          "diet"         
"year"          "site.age.cent" "status.focals" "N"             "i"            
"site.age.init" "site.age"    


## Preliminaries

### Load cleaned, merged and scrubbed data

```{r}
load(file = "./data/ann_counts.RData") #loads "ann_counts"
```


## Load libraries

Function that calls all libraries needed for analysis
```{r}
source("./R/load_libraries.R")
load_libraries()
```

## Look at data

```{r}
names(ann_counts)
```


## Index for Removing aceitillar

Aceitillar is run seperately using 1-way ANOVA style models in the next script in this series.

```{r}
i.N.aceit <- which(ann_counts$site == "Aceitillar")
```



## Exploratory graphs

These graphs treat age as a continuous variable and show the non-linearites that are ignored in the regression-style GLMM anlaysis.  

```{r}
i1 <- which(ann_counts$status.focals == "mig")
i2 <- which(ann_counts$status.focals == "res")
ggplot(data = ann_counts[i2,],
       aes(y = N,
           x = site.age,
           color = status.focals)) +
  facet_wrap(~spp.code,scale = "free") +
  geom_point() +
  geom_smooth(se = F)
```

### Contrast modeling: All species

#### Focal species

Species with sparcer counts are more difficult to model b/c they may not occur at all sites or may be very rare in some years.  

```{r}
focal.mig.x2 <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW",
               "MAWA"#,
               #"NOPA",
               #"PALM"
               ) #add to poisson

focal.res.x2 <- c( #"HLCU","HILC",         #fix - which name to use?
                #"STOF",
                "RLTH",
               "NOMO",
                #"GRWA",     
                "HILC",
                "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "STOF",
               "GABU",
               "BWVI"#, #32; Aceitillar / Cueva +misc
               #"GREP"
               ) #18 just Aceitillar - ever yr

focals.x2 <- c(focal.mig.x2,focal.res.x2)
```


### Loop over each focal spp

Note: use all years; don't have to drop due to overlap

```{r}
#focal species to use
spp.to.use.x2 <- focal.mig.x2

#create list
## models w/time trend
x.model.list <- as.list(spp.to.use.x2)
names(x.model.list) <- spp.to.use.x2

## models w/time trend + time^2
x2.model.list <- as.list(spp.to.use.x2)
names(x2.model.list) <- spp.to.use.x2

options(warn = -1)
```




### Run models
```{r}
for(i in 1:length(spp.to.use.x2)){
  
i.spp <- which(ann_counts$spp.code == spp.to.use.x2[i])

print(i)
print(spp.to.use.x2[i])
dim(ann_counts[i.spp, ])
m.x.spp <- bglmer(N ~ 1 + site.age.cent +
               (1|i) + 
               (1|year)+
                 (1|site),
               offset = log(net.hours),
              data = ann_counts[i.spp, ],
              family = poisson ,
              glmerControl(optimizer = "Nelder_Mead")
              )
m.x2.spp <- update(m.x.spp, . ~ . + I(site.age.cent^2))

x.model.list[[i]] <- m.x.spp
x2.model.list[[i]] <- m.x2.spp

}

```

#compare models

```{r}
likrat_test_stats <- function(m1,m2){
  mod.out <- anova(m1,m2)
  test.stats <- c(paste(mod.out$Df,collapse = ",")
  ,round(mod.out$Chisq[2],4)
  ,round(mod.out$"Pr(>Chisq)"[2],4))
  
  names(test.stats) <- c("df","Chisq","p")
  return(test.stats)
}
```

```{r}
x2.teststat.list <- x.model.list
for(i in 1:length(x.model.list)){
  x2.teststat.list[[i]] <- likrat_test_stats(x.model.list[[i]],
                                            x2.model.list[[i]])
}

```

```{r}
library(data.table)
x2.teststat.df <- data.frame(t(as.data.table(x2.teststat.list,keep.rownames = T)))

names(x2.teststat.df) <- c("df","chi2","p")

x2.teststat.df <- cbind(Species = row.names(x2.teststat.df),
      x2.teststat.df)
x2.teststat.df$p <- as.numeric(as.character(x2.teststat.df$p))
x2.teststat.df$sig <- ""
x2.teststat.df$sig <- ifelse(x2.teststat.df$p < 0.1, ".",x2.teststat.df$sig)
x2.teststat.df$sig <- ifelse(x2.teststat.df$p < 0.05, "*",x2.teststat.df$sig)
x2.teststat.df$sig <- ifelse(x2.teststat.df$p < 0.01, "**",x2.teststat.df$sig)
```



Most abundant site
```{r}
x2.teststat.df$most.abundant <-
  c("Morelia",#OVEN
    "Corral/Aceitillar", #BAWW
    "Caoba", #COYE
    "Caoba", #AMRE
    "Caoba", #CMWA
    "Caoba", #BTBW
    "Cueva", #PAWA
    "Cueva/Caoba", #PRAW
    "" #MAWA
    )
```



## Save output

```{r}
write.csv(x2.teststat.df, file = "./tables/year_squared_test_stats.csv")
```

