---
title: "Untitled"
author: "Nathan Brouwer"
date: "January 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

Note that for aceitillar sampling only ever occured in Nov, Jan and March, and that this only occured in 1996-1997, and 1997-1998.  In 1999 and beyond all sampling occurred in January (Or only january data considered).  

Note that sampling in eg January 1999 is coded as "1998-1999" for consistency witht he prior years and the Mencia data.

In contrast, Mencia sampling occurred in Nov, Jan and Feb (never March) in all years except 2008 (final year), when no sampling is indicate in March.



2004 2005 2006 2007 2008            #5 yrs @ mencia
1997 1998 1998 2000 2001 2002 2003  #7 yrs @ aceitilaR


## Load Data

### Load Mencia data

```{r}
# duplicated code?


# filename <- "mencia_site_persist_clean.csv"
# filename <- here::here("data-raw",
#            "data-raw-mencia",
#            "mencia_site_persistence",
#            filename)
# mencia_site_persist <- read.csv(filename)
```


### Load Mencia data

```{r}
filename <- "mencia_site_persist_clean.csv"
filename <- here::here("data-raw",
           "data-raw-mencia",
           "mencia_site_persistence",
           filename)
mencia_site_persist <- read.csv(filename, stringsAsFactors = F)
```

### Load Aceitillar data
```{r}
filename <- "aceit_site_persist_clean.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)
aceit_site_persist <- read.csv(filename, stringsAsFactors = F)
```

### Merge just for overall "site persistence" rating

this ranking uses all years of data for both sites, even though the sites differ in sampling effort both in terms of resight  effort (less resight for Aceit) AND number of years (more years Aceit).  Therefor it is a very naive categorications
```{r}
col.names <- intersect(names(aceit_site_persist),
        names(mencia_site_persist))

```

```{r}
site_persist <- merge(aceit_site_persist[,col.names],
      mencia_site_persist[,col.names], all = TRUE)
```

Check size

```{r}
dim(aceit_site_persist)
dim(mencia_site_persist)
dim(site_persist)
```


## save .RData files

saved to "data" folder where working data for real analyses is stored

```{r}
filename <- "site_persist.RData"
filename <-here::here("data",
            filename)
save(site_persist, 
           file = filename,row.names = F)
```

```{r}

```

