---
title: "Finalize setup of data"
author: "Nathan Brouwer"
date: "January 30, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

This script produces processed dataframe that are used directly for analyses.  It produces 


### Dataframes produced

#### Individual-level data
* ann_caps  = original dataframe: all data; includes recaptures between years (recaptures _within_ years are removed earlier (?))
* ann_caps4age_ratio = just focal spp; rare species removed
* ann_caps_1st_captures = just 1st captures of each individual; recaptures in subsequent years removed; NOT USED for analysis
* ann_caps4sex_ratio = subset of birds than can be reliably sexed;

#### Species level data
* ann_counts = total number of each spp observed at each site-year combo

#### Community data

* spp_matrix = number of individuals of each spp captured per year, wide format
* community_dat = spp richness, spp diversity etc by year
* community_dat_pooled = data for each site pooled accross years


## Preliminaries


### Libraries
```{r}
library(reshape2)
```


### Load cleaned, merged and scrubbed data

```{r}
load("./data/ann_caps.RData")
```


### Remove any "LOST" spp code

```{r}
ann_caps <- ann_caps[-which(ann_caps$spp.code == "LOST"), ]
```



### Set up factors for easy interp with summary()

```{r}
ann_caps$site <- factor(ann_caps$site,
                        levels = c("La Cueva","La Caoba",
                                  "Morelia","El Corral", "Aceitillar"))
#ann_caps$spp.code <- factor(ann_caps$spp.code)

ann_caps$sex <- factor(ann_caps$sex,
                            levels = c("U","F","M"))

ann_caps$status.focals <- factor(ann_caps$status.focals)

```


### Look at data
```{r}
summary(ann_caps[, c("spp.code","site",
                     "sex","age.AHY.01",
                     "site.age","site.age.init",
                     "status.focals")])


#with(ann_caps, table(site,site.age,useNA = "always"))

with(ann_caps, table(age,age.AHY.01,useNA = "always"))
```

## Fix typos

```{r}
#Stolid flycatch didn't have diet info
ann_caps[which(ann_caps$spp.code == "STOF"),c("diet")]<- "O"

#a few typos for ZEND
## add diet and habitat




```


## Set up dataframes for particular analyeses


### Analysis 1: Set up dataframe for age ratio analysis: 

For the age ratio data 

* non-focal species need to be removed and 
* any recaptures of that individual w/in a year need to be removed

Further set up is done for particular analyes within the appropriate scripts (eg, remove aceitillar, remove overlap of youngest pasture sites for ordered ANOVA analysis)

#### Focal spp for age ratio analysis

These are species w/ sufficent samples sizes / balance
```{r}
# focal migrants
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

# focal residents
focal.res <- c("HLCU","HILC",         #fix - which name to use?
               "STOF",
               "RLTH","NOMO",
               "GRWA",       "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "GABU",
               "BWVI", #32; Aceitillar / Cueva +misc
               "GREP") #18 just Aceitillar - ever yr

focals <- c(focal.mig,
            focal.res)
```


Update "status focals" column
```{r}
i.res <- which(ann_caps$spp.code %in% focal.res)

ann_caps$status.focals[i.res] <- "res"
```



#### Subset by focal spp

```{r}
i.focals <- which(ann_caps$spp.code %in% focals)
 
ann_caps4age_ratio <- ann_caps[i.focals,]
```

```{r}
summary(factor(ann_caps4age_ratio$spp.code))
```

#### Which spp are are and are not "focal"

```{r}
summary(factor(as.character(ann_caps[-i.focals,"spp.code"])))
```


```{r}
summary(factor(as.character(ann_caps[i.focals,"spp.code"])))
```




#### Isolate unique individuals

* remove repeated measures; this simplifies the models but tosses out data.  
* This was not ultimately used and is left for reference

```{r}

i.unique <- match(unique(ann_caps4age_ratio$band),
                  ann_caps4age_ratio$band)

dim(ann_caps4age_ratio)
length(unique(ann_caps4age_ratio$band))
length(i.unique)


ann_caps_1st_captures <- ann_caps4age_ratio[i.unique,]
```



### Analysis 2: Set up for sex ratio analysis: Isoalte spp that can be sexed reliably

Note: using ann_caps not ann_caps4age_ratio just to be safe


#### Number of individuals per sex

```{r}
dcast(data = ann_caps,
      formula = spp.code ~ sex,
      fun.aggregate = length,
      value.var = "age.AHY.01")
```

Generate indices for birds that can be reliablly sexed
```{r}
spp.sex <- c("AMRE","BAWW","BFGR","BTBW",
             "CMWA","COYE","GABU","PRAW",
             "YFGR")

#remove rarely sexed birds
i.use <- which(ann_caps$spp.code %in% spp.sex)

```

Check resulting datafrmaes
```{r}
#spp to use
dcast(data = ann_caps[i.use,],
      formula = spp.code ~ sex,
      fun.aggregate = length,
      value.var = "age.AHY.01")

#spp dropped
dcast(data = ann_caps[-i.use,],
      formula = spp.code ~ sex,
      fun.aggregate = length,
      value.var = "age.AHY.01")
```

#### Subset data
Note: using ann_caps not ann_caps4age_ratio just to be safe
```{r}
ann_caps4sex_ratio <- ann_caps[i.use,]
```



### Analysis 3: Reshape data to tabulate annual counts

Annual number of each species captured.  Data is no longer at individual level!

```{r}
tot.obs.by.hab <- dcast(data = ann_caps,
      formula = spp.code  ~  site,
      value.var = "spp.code",
      fun.aggregate = length)

tot.obs <- dcast(data = ann_caps,
      formula = spp.code  ~  .,
      value.var = "spp.code",
      fun.aggregate = length)
names(tot.obs) <- gsub("^\\.","N",names(tot.obs))

dim(tot.obs[which(tot.obs$N < 5), ])
dim(tot.obs[which(tot.obs$N < 6), ])
dim(tot.obs[which(tot.obs$N < 7), ])
dim(tot.obs[which(tot.obs$N < 10), ])
dim(tot.obs[which(tot.obs$N < 15), ])
dim(tot.obs[which(tot.obs$N < 20), ])

n.min <- 15

rarespp <- tot.obs[which(tot.obs$N < n.min), "spp.code"]

```

```{r}
i.rare <- which(ann_caps$spp.code %in% rarespp)
tot.obs.by.site <- dcast(data = ann_caps[-i.rare,],
      formula = spp.code  ~  site,
      value.var = "spp.code",
      fun.aggregate = length)
names(tot.obs.by.site) <- gsub("^\\.","N",names(tot.obs.by.site))

```


####
```{r}
summary(factor(ann_caps$year))
```


#### Reshape for annual data
```{r}
ann_counts <- dcast(data = ann_caps[,],
      formula = year +year.num + site + site.age.cent + 
        spp.code + status.focals + hab1 + diet ~ .,
      value.var = "spp.code",
      fun.aggregate = length)

names(ann_counts) <- gsub("^\\.","N",names(ann_counts))

summary(factor(ann_counts[-which(ann_counts$status.focals %in% c("mig","res")), "spp.code"]))

i.newfoc <- which(ann_counts$spp.code %in% c("MAWA","NOPA","PALM"))
ann_counts[i.newfoc,"status.focals"] <- "mig"

ann_counts$status.focals[is.na(ann_counts$status.focals)] <- "res"

summary(factor(ann_counts$status.focals))
```



```{r}
summary(factor(ann_counts$year))
```



#### Add zeros

Make a dataframe with all spp-site-year combos
```{r}

#mencia sub df
mencia.expand <- with(ann_counts, 
                     expand.grid(spp.code = unique(spp.code),
                             site = c("El Corral", "La Caoba", 
                                      "La Cueva", "Morelia"),
                             year.num = 2003:2007)) 
mencia.expand$year <- paste(mencia.expand$year,
                            mencia.expand$year+1,
                            sep = "-")
#aceit subdf
aceit.expand <- with(ann_counts, 
                     expand.grid(spp.code = unique(spp.code),
                             site = c("Aceitillar"),
                             year.num = 1996:2001))


aceit.expand$year <- paste(aceit.expand$year,
                            aceit.expand$year+1,
                            sep = "-")

#combine
dat.expand <-rbind(aceit.expand,
                   mencia.expand)

```


#### Add spp info to dat.expand 

load spp info and subset by spp acutally in study
```{r}

spp.list <- read.csv(file = "./data/spp_list.csv", stringsAsFactors = F)


summary(factor(spp.list$hab1))
summary(factor(spp.list$diet))

#remove spp not used in study
i.spp.use <- which(spp.list$spp.code %in% unique(ann_counts$spp.code))

spp.list2 <- spp.list[i.spp.use,]

dim(spp.list)
dim(spp.list2)
```

Merge spp list and dat.expand
```{r}
dat.expand2 <- merge(dat.expand,
                     spp.list2[,c("spp.code","hab1","diet")], all = TRUE)

dim(dat.expand)
dim(dat.expand2)

```


#### Merge dat.expand2 and count data
```{r}
ann_counts2 <- merge(ann_counts, 
                     dat.expand2, all = TRUE)                            


dim(dat.expand)
dim(dat.expand2)                             
dim(ann_counts)
dim(ann_counts2)


```


#### Remove rare species

```{r}
i.rare <- which(ann_counts2$spp.code %in% rarespp)
length(rarespp)
length(i.rare)

ann_counts3 <- ann_counts2[-i.rare,]
dim(ann_counts3)
```


```{r}
ann_counts <- ann_counts3




```

### Check - NAs get introduced for year

All of these correspond to valeus of 0 for abundance (set below).
So, year-bird-xxx combos w/o any captures are getting set to NA?
```{r}
summary(factor(ann_counts$year))
with(ann_counts, table(year,site,useNA = "always"))
ann_counts[which(is.na(ann_counts$year) == TRUE), ]
```



```{r}
# focal migrants
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW",
               "MAWA","NOPA","PALM")

# focal residents
focal.res <- c("HLCU","HILC",         #fix - which name to use?
               "STOF",
               "RLTH","NOMO",
               "GRWA",       "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "GABU",
               "BWVI", #32; Aceitillar / Cueva +misc
               "GREP") #18 just A
```


```{r}
ann_counts$status.focals[which(ann_counts$spp.code %in% focal.res)] <- "res"
ann_counts$status.focals[which(ann_counts$spp.code %in% focal.mig)] <- "mig"

ann_counts$status.focals[is.na(ann_counts$status.focals)] <- "res"

```

Set NAs to zero
```{r}
ann_counts$N[is.na(ann_counts$N)] <- 0
```

#### Add Observation level random effect

```{r}
ann_counts$i <- factor(1:dim(ann_counts)[1])
```


#### Calculate age of sites

Because the study occurs over several years the sites undergo succession and their exact age each year is importnat to consider

```{r}
ann_counts$site.age.init <- NA
ann_counts$site.age.init[which(ann_counts$site == "La Cueva")] <- 2
ann_counts$site.age.init[which(ann_counts$site == "La Caoba")] <- 5
ann_counts$site.age.init[which(ann_counts$site == "Morelia")]  <- 10
ann_counts$site.age.init[which(ann_counts$site == "El Corral")]  <- 20

 
ann_counts$site.age <- ann_counts$site.age.init + 
                      ann_counts$year.num-2003

ann_counts$site.age.cent <- scale(ann_counts$site.age, scale = F)
```


### Merge with net hours


#### Load net hours

```{r}
net_hrs <- read.csv(file = "./data/net_hours.csv")[,1:5]
```


```{r}
ann_counts4 <- merge(ann_counts,net_hrs,all = T)
dim(ann_counts) == dim(ann_counts4)


```


```{r}
ann_counts <- ann_counts4
summary(factor(ann_counts$year))
```



### Analysis 4: Community analyses by year

#### Reshape for general community analyses
```{r}
library(vegan)

spp_matrix <- dcast(data = ann_caps,
      formula = site + year + 
                year.num + site.age + 
                site.age.init + 
                site.age.cent ~ spp.code,
      value.var = "spp.code",
        fun.aggregate = length)

#isoalte colnames that are spp
i.spp.names <- grep("[A-Z][A-Z][A-Z][A-Z]",names(spp_matrix))

with(spp_matrix,table(site, site.age))

#isolate predictoras
community_dat <- spp_matrix[,-i.spp.names]

community_dat$simpson <- diversity(spp_matrix[,i.spp.names],index = "shannon")

convert2.zero <- function(x){ifelse(x > 0,1,0)}

temp <- spp_matrix[,i.spp.names]
temp <- apply(temp,2,convert2.zero)

specnumber(temp)

community_dat$spp.rich <- rowSums(temp)

```


##### Calculate evenness

Pilou's evenness
```{r}
community_dat$evenness <- community_dat$simpson/log(community_dat$spp.rich)

```

##### Calculate dominance 

?


#### Merge with nethours

```{r}
community_dat2 <- merge(community_dat,net_hrs,all = T)
dim(community_dat2)
dim(community_dat)

dim(community_dat2) == dim(community_dat)
summary(community_dat2$site)
summary(factor(community_dat2$year))
```












### Analysis 5: Community analyses by site (pooling years)

Pool data accross years w/in a site

#### Reshape for general community analyses
```{r}
library(vegan)

spp_matrix2 <- dcast(data = ann_caps,
      formula = site + site.age.init ~ spp.code,
      value.var = "spp.code",
        fun.aggregate = length)

#isoalte colnames that are spp
i.spp.names2 <- grep("[A-Z][A-Z][A-Z][A-Z]",names(spp_matrix2))



#isolate predictoras
#isolate predictoras
community_dat2 <- spp_matrix2[,-i.spp.names2]

community_dat2$simpson <- diversity(spp_matrix2[,i.spp.names2],index = "shannon")

convert2.zero <- function(x){ifelse(x > 0,1,0)}

temp <- spp_matrix2[,i.spp.names2]
temp <- apply(temp,2,convert2.zero)

specnumber(temp)

community_dat2$spp.rich <- rowSums(temp)


```


##### Calculate evenbess
```{r}
community_dat2$evenness <- community_dat2$simpson/log(community_dat2$spp.rich)
```


Rename
```{r}
spp_matrix_pooled <- spp_matrix2
community_dat_pooled <- community_dat2
```









### Analysis 6: Species characteristics / traits


#### Check for any missing traits 

```{r}
(ann_caps[is.na(ann_caps$hab1), c("spp.code","site")])
(ann_caps[is.na(ann_caps$diet), c("spp.code","site")])

```


#### Reshape by traits 

```{r}
summary(factor(ann_caps$hab1))
ann_caps$hab1 <- as.character(ann_caps$hab1)
ann_caps$hab1 <- gsub("D$","DF",ann_caps$hab1)
summary(factor(ann_caps$hab1 ))
hab_matrix <- dcast(data = ann_caps,
      formula = site + year + 
                year.num + site.age + 
                site.age.init + 
                site.age.cent ~ hab1,
      value.var = "spp.code",
        fun.aggregate = length)
#names(hab_matrix)[dim(hab_matrix)[2]] <- "hab.NA"
hab.tot <- apply(hab_matrix[,c("DF","EF","SF")],1,sum)


diet_matrix <- dcast(data = ann_caps,
      formula = site + year + 
                year.num + site.age + 
                site.age.init + 
                site.age.cent ~ diet,
      value.var = "spp.code",
        fun.aggregate = length)
#names(diet_matrix)[dim(diet_matrix)[2]] <- "diet.NA"

diets <- c("C","F","G","I","N","O")
diet.tot <- apply(diet_matrix[,diets],1,sum)

trait_dat <- cbind(hab_matrix,
                   diet_matrix[,diets])

#check that row totals match
plot(diet.tot , hab.tot)

habs <- c("DF","EF","SF")
names(trait_dat)[which(names(trait_dat) %in% habs)] <- paste("hab",habs,sep = ".")
names(trait_dat)[which(names(trait_dat) %in% diets)] <- paste("diet",diets,sep = ".")


trait_dat$tot <- diet.tot
```







## Dataframes summary

These are the different datasets generate by this code

### Dataframes

#### Individual-level data
* ann_caps  = original dataframe: all data; includes recaptures between years (recaptures _within_ years are removed earlier (?))
* ann_caps4age_ratio = just focal spp; rare species removed
* ann_caps_1st_captures = just 1st captures of each individual; recaptures in subsequent years removed; NOT USED for analysis
* ann_caps4sex_ratio = subset of birds than can be reliably sexed;

#### Species level data
* ann_counts = total number of each spp observed at each site-year combo

#### Community data

* spp_matrix = number of individuals of each spp captured per year, wide format
* community_dat = spp richness, spp diversity etc by year
* community_dat_pooled = data for each site pooled accross years
* trait_dat = compositional data on habitat abd diet
### Save new dataframes 

Original dataframe is ann_caps.  

```{r}
#age ratio
save(ann_caps4age_ratio, file = "./data/ann_caps4age_ratio.RData")

#sex ratio
save(ann_caps4sex_ratio, file = "./data/ann_caps4sex_ratio.RData")

#counts of each spp per year
summary(factor(ann_counts$year))
save(ann_counts, file = "./data/ann_counts.RData")


# community data
## wide matrix of spp by year
save(spp_matrix, file = "./data/spp_matrix.RData")

## wide matrix pooled
save(spp_matrix_pooled, file = "./data/spp_matrix_pooled.RData")

##richness, eveness, div
save(community_dat, file = "./data/community_dat.RData")

##richness, eveness, div by site

save(community_dat_pooled, file = "./data/community_dat_pooled.RData")


## trait data

save(trait_dat, file = "./data/trait_dat.RData")

```



```{r}

```




```{r}
x <- dcast(data = ann_caps,
           formula = spp.code ~ site,
           value.var = "spp.code",
           fun.aggregate = length)
write.csv(x, file = "temp.csv")
```

