---
title: "Calculate Condition Indicies"
author: "Nathan Brouwer"
date: "February 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Calcualte Scaled mass indiex m.hat
as per Peig & Green 2009 "New perspectives for estimating body condition from mass/length data: the scaled mass index as an alternative method" Oikos 118: 1883-1891


## Load Libraries
```{r}
library(ggplot2)
library(ggpubr)
library(stringr)
```


## Load data

```{r}
load(file = here::here("data-raw",
                                  "data-raw-bodymass",
                                  "condition.RData"))
```


## Prep data
```{r}
condition$wing.log <- log(condition$wing)
condition$mass.log <- log(condition$mass)
```


## Extract year

```{r}
condition$year
```



## Focal species

```{r}
# focal migrants
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW",
               "MAWA","NOPA","PALM")

# focal residents
focal.res <- c("HLCU","HILC",         #fix - which name to use?
               "STOF",
               "RLTH","NOMO",
               "GRWA",       "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "GABU",
               "BWVI", #32; Aceitillar / Cueva +misc
               "GREP") #18 just A

focals <- c(focal.mig,focal.res)
```



## Look at spp-site combos

```{r}
i.focs <- which(condition$spp.code %in% focals)

spp.with.condition <- dcast(data = condition[i.focs,],
      formula = spp.code ~ site,
      value.var = "wing")

```


Function to calculate M.hat
```{r}
M.hat <- function(Mi, Li, L0,b.SMA){
  Mi*((L0/Li)^b.SMA)
}
```

## Run sma regression

Major axis regression
```{r}
condition2 <- condition
condition2$M.hat.i <- NA
condition2$e.i <- NA



condition2$ratio <- with(condition2, mass/wing)

x1 <- run_mhat_loop(spp = focals,
        dat = condition2,
        L0 = "mean")
x2 <- run_mhat_loop(spp = focals,
        dat = condition2,
        L0 = "")


x1$params
summary(x1$condition$M.hat.i)
summary(x2$condition$M.hat.i)
head(x2$condition)
condition2 <- x2$condition

plot(x1$condition$M.hat.i,
     x2$condition$M.hat.i)
```



## ggpubr
```{r}
library(ggpubr)

condition2$site.age <- factor(condition2$site.age,
                              levels = c("2","5","10","20","mature"))

#make new sex variable
condition2$sex2 <- as.character(condition2$sex)

#change all NAs to "Unknown"
condition2$sex2[which(is.na(condition2$sex2) == TRUE)] <- "Unknown"

# with(condition2,
#      table(sex, sex2,useNA = "always"))
# with(condition2,
#      table(sex, sex2,spp.code,useNA = "always"))

#Ovenbird; change rare "F" to unknown
condition2$sex2[which(condition2$sex2 == "F" &
        condition2$spp.code == "OVEN")] <- "Unknown"

condition2$sex2 <- ifelse(condition2$sex2 == "Unknown" &
        condition2$spp.code == "CMWA",
        NA,condition2$sex2)

condition2$sex2 <- ifelse(condition2$sex2 == "Unknown" &
        condition2$spp.code == "AMRE",
        NA,condition2$sex2)

condition2$sex2 <- ifelse(condition2$sex2 == "Unknown" &
        condition2$spp.code == "BFGR",
        NA,condition2$sex2)
condition2$sex2 <- ifelse(condition2$sex2 == "Unknown" &
        condition2$spp.code == "YFGR",
        NA,condition2$sex2)

condition2$sex2 <- ifelse(condition2$sex2 == "Unknown" &
        condition2$spp.code == "GABU",
        NA,condition2$sex2)

with(condition2,
     table(sex, sex2,useNA = "always"))





```


## Save

```{r}
condition.Mhat <- condition2
save(condition.Mhat,file = "./data/condition_Mhat.RData")
```





### Exploratory plots

Subset by status; remove bad entries
```{r}
i.mig <- which(condition2$stat.focals == "mig" &
              is.na(condition2$sex2) == FALSE)
i.res <- which(condition2$stat.focals == "res" &
              is.na(condition2$sex2) == FALSE)

i.all <- c(i.mig,i.res)
i.BANA.bad <- with(condition2[i.all,],
     which(spp.code == "BANA" & sex2 == "M"))
i.BAWW.bad <- with(condition2[i.all,],
     which(spp.code == "BAWW" & sex2 == "Unknown"))

i.all <- i.all[-which(i.all %in% c(i.BANA.bad,i.BAWW.bad)) ]
```


Plot residents
```{r}
#Plot
ggerrorplot(data = condition2[i.res,],
            y = "M.hat.i",
            x = "site.age",
            desc_stat = "mean_se",
            color = "sex2") + facet_wrap(~spp.code,
                                        scale = "free")
```

Plot All
```{r}
#Plot
ggerrorplot(data = condition2[i.all,],
            y = "M.hat.i",
            x = "site.age",
            desc_stat = "mean_se",
            color = "sex2") + 
             facet_wrap(~spp.code,
                          scale = "free")
```




