---
title: "Untitled"
author: "Nathan Brouwer"
date: "January 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

Note that for aceitillar sampling only ever occured in Nov, Jan and March, and that this only occured in 1996-1997, and 1997-1998.  In 1999 and beyond all sampling occurred in January.  Note that sampling in January 1999 is coded as "1998-1999" for consistency witht he prior years and the Mencia data.

In contrast, Mencia sampling occurred in Nov, Jan and Feb (never March) in all years except 2008, when not sampling is indicate in March.


## Load Data

### Load Mencia data

```{r}
filename <- "mencia_site_persist_clean.csv"
filename <- here::here("data-raw",
           "data-raw-mencia",
           "mencia_site_persistence",
           filename)
mencia_site_persist <- read.csv(filename)
```

```{r}
list.files(path = here::here("data-raw",
           "data-raw-mencia",
           "mencia_site_peristence"))
```


### Load Mencia data

```{r}
filename <- "mencia_site_persist_clean.csv"
filename <- here::here("data-raw",
           "data-raw-mencia",
           "mencia_site_persistence",
           filename)
mencia_site_persist <- read.csv(filename, stringsAsFactors = F)
```

### Load Aceitillar data
```{r}
filename <- "aceit_site_persist_clean.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)
aceit_site_persist <- read.csv(filename, stringsAsFactors = F)
```

### Merge
```{r}
col.names <- intersect(names(aceit_site_persist),
        names(mencia_site_persist))

```

```{r}
site_persist <- merge(aceit_site_persist[,col.names],
      mencia_site_persist[,col.names], all = TRUE)
```

Check size

```{r}
dim(aceit_site_persist)
dim(mencia_site_persist)
dim(site_persist)
```


## save .RData files

saved to "data" folder where working data for real analyses is stored

```{r}
filename <- "site_persist.RData"
filename <-here::here("data",
            filename)
save(site_persist, 
           file = filename,row.names = F)
```

```{r}

```

