---
title: "Untitled"
author: "Nathan Brouwer"
date: "February 22, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



NOTE: No new captures are listed in teh site-persistence dataframe for Nov 2007 or Jan 2008 b/c 2008 was that last year of the study

However, New captures from Nov should be included b/c they could be recaptured in Nov 2007 b/c they could be recapped in Jan 2008


NOTE: not resights entered for Nov 2007



## Load Mencia site persistence data




```{r}
library(here)


fi <- here::here("data-raw",
           "data-raw-mencia",
           "mencia_site_persistence",
           "mencia_site_persist_clean.csv")
mencia.sp <- read.csv(fi)
```



#look at columns related to 1st capture
```{r}
date.1st.cap <- c("date.cap.orig","yr","mo","day")
head(mencia.sp[,date.1st.cap])
```


#
Construct names of columns w/t capture/ resight info
```{r}
col.1st.cap <- with(mencia.sp,paste("rc",yr,mo,sep = "."))
col.1st.cap <- gsub("20","",col.1st.cap)
col.1st.cap <- gsub(".12$",".J",col.1st.cap) #captures in Dec are Nominally JANUARY
col.1st.cap <- gsub(".11$",".N",col.1st.cap)
col.1st.cap <- gsub(".1$",".J",col.1st.cap)
col.1st.cap <- gsub(".2$",".F",col.1st.cap)
col.1st.cap <- gsub(".3$",".F",col.1st.cap) #captures in March are nominally in FEB


#add column to main datafram
mencia.sp$yr.mo.1st.cap <-col.1st.cap
mencia.sp$yr.mo.1st.cap <- gsub("rc.","",mencia.sp$yr.mo.1st.cap)



```


Look at data
```{r}
head(mencia.sp[,unique(col.1st.cap)],100)


levels(factor(col.1st.cap))
levels(factor(col.1st.cap))


names(mencia.sp)[grep("^rc",names(mencia.sp))][grep("08",names(mencia.sp)[grep("^rc",names(mencia.sp))])]

setdiff(names(mencia.sp)[grep("^rc",names(mencia.sp))],
        levels(factor(col.1st.cap)))
```



## Code 1st time captured
```{r}
length(col.1st.cap)

for(i in 1:length(col.1st.cap)){
 # print(i)
  
  j <- which(names(mencia.sp) == col.1st.cap[i])
  
  #print(j)
  
  mencia.sp[,j] <- ifelse(mencia.sp[,j] == "bf","-5",mencia.sp[,j])
  mencia.sp[,j] <- as.numeric(mencia.sp[,j])
  
   mencia.sp[i,j] <- -1
  
   
}
```



Check that it worked

```{r}
length(which(mencia.sp$rc.03.N == -1))
length(which(mencia.sp$yr == 2003 & mencia.sp$mo == 11))
```



```{r}
summary(mencia.sp[,grep("^rc",names(mencia.sp))])
```



## Code recaps as Y/N

```{r}
mencia.sp$recapYN <- ifelse(mencia.sp$num.rc.rs >0, "sp","not.sp")
mencia.sp$recapYN  <- factor(mencia.sp$recapYN )
```


##

```{r}
i.1st.cap.03.N <- which(mencia.sp$rc.03.N == -1)


```

## Look at ratio of caps to recaps
```{r}
with(mencia.sp[i.1st.cap.03.N,], table(spp.code, recapYN) )
```


```{r}
mencia.sp$site <- as.character(mencia.sp$site)
mencia.sp$site <- factor(mencia.sp$site,
                         levels = c("La Cueva","La Caoba","Morelia","El Corral", "Aceitillar"))
summary(mencia.sp$site)

mencia.sp$spp.code[which(mencia.sp$spp.code == "BANA ?")] <- "BANA"
mencia.sp$stat  <- as.character(mencia.sp$stat )
mencia.sp$stat.MvsR <- ifelse(mencia.sp$stat == "RE","R",mencia.sp$stat)

```


```{r}
summary(mencia.sp$stat)
```


```{r}
i.1st.cap.03.N <- which(mencia.sp$rc.03.N == -1)
i.1st.cap.03.all <- which(mencia.sp$rc.03.N == -1 |
                          mencia.sp$rc.04.J == -1 |
                          mencia.sp$rc.04.F == -1)
length(i.1st.cap.03.N)
length(i.1st.cap.03.all)

m.sp.mencia <- bglmer(recapYN ~ site*stat.MvsR +
                        (1|spp.code) +
                        (1|yr.mo.1st.cap), 
                   family = binomial,
                   data = mencia.sp[i.1st.cap.03.all,])
summary(m.sp.mencia)

m.sp.mencia.means <- bglmer(recapYN ~ -1 + site:stat.MvsR + 
                        (1|spp.code), 
                   family = binomial,
                   data = mencia.sp[i.1st.cap.03.all,])
summary(m.sp.mencia.means)
```




