---
title: "Untitled"
author: "Nathan Brouwer"
date: "February 22, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##


## Load Mencia site persistence data




```{r}
library(here)


fi <- here::here("data-raw",
           "data-raw-aceitillar_96-02",
           "aceit_site_persist_clean.csv")
aceitillar.sp <- read.csv(fi)
```


## Create month columns

```{r}
#aceitillar.sp$date.cap.orig
aceitillar.sp$yr <- lubridate::year(aceitillar.sp$date.cap.orig)
aceitillar.sp$mo <- lubridate::month(aceitillar.sp$date.cap.orig)
aceitillar.sp$mo2 <- lubridate::month(aceitillar.sp$date.cap.orig,
                                      label = T,
                                      abbr = T)
aceitillar.sp$mo2 <- gsub("Jan","J",aceitillar.sp$mo2)
aceitillar.sp$mo2 <- gsub("Nov","N",aceitillar.sp$mo2)
aceitillar.sp$mo2 <- gsub("Mar","M",aceitillar.sp$mo2)
aceitillar.sp$mo2 <- gsub("Feb","M",aceitillar.sp$mo2) #Feb is nominally Mach

summary(factor(aceitillar.sp$mo2))
```



## Construct names of columns w/t capture/ resight info
```{r}
aceit.col.1st.cap <- with(aceitillar.sp,paste(mo2,yr,sep = "."))

aceit.col.1st.cap <- gsub("20","",aceit.col.1st.cap)
aceit.col.1st.cap <- gsub("19","",aceit.col.1st.cap)


#add column to main datafram
aceitillar.sp$yr.mo.1st.cap <-aceit.col.1st.cap
```




# Note
There are 132 records of  birds 1st capture in  birds 1st capture in   march of 2001  in the sp dataframe no march records for 99, 00, or 02 as  originally implied  by data structure

```{r}
which(aceitillar.sp$yr.mo.1st.cap == "M.99")
which(aceitillar.sp$yr.mo.1st.cap == "M.00")
length(which(aceitillar.sp$yr.mo.1st.cap == "M.01")) #132
which(aceitillar.sp$yr.mo.1st.cap == "M.02")

```



# Convert alpha codes to nmber

```{r}
j.cols <- grep("[NDJFM]",names(aceitillar.sp))
for(i in 1:length(j.cols)){
 aceitillar.sp[,j.cols[i]]  <- as.character(aceitillar.sp[,j.cols[i]]  )
  aceitillar.sp[,j.cols[i]] <- gsub("C","1",aceitillar.sp[,j.cols[i]])
  aceitillar.sp[,j.cols[i]] <- gsub("CS","1",aceitillar.sp[,j.cols[i]])
  aceitillar.sp[,j.cols[i]] <- gsub("CC","1",aceitillar.sp[,j.cols[i]])
  aceitillar.sp[,j.cols[i]] <- as.numeric(aceitillar.sp[,j.cols[i]])
}

summary(aceitillar.sp[,j.cols])
head(aceitillar.sp[,j.cols])
```


## Code 1st time captured
```{r}
length(col.1st.cap)

for(i in 1:length(aceit.col.1st.cap)){
 # print(i)
  
  #March 2001 birds - skip
  if(aceit.col.1st.cap[i] == "M.01"){next} 
  j <- which(names(aceitillar.sp) == aceit.col.1st.cap[i])
  
  #print(j)
  
  
  aceitillar.sp[i,j] <- -1
  
   
}
```

Check that it worked

```{r}
length(which(aceitillar.sp$N.96 == -1))
length(which(aceitillar.sp$yr == 1996 & aceitillar.sp$mo == 11))
head(aceitillar.sp[,j.cols])
```



## Determine site persistence

```{r}
aceitillar.sp$recapYN <- NA
aceitillar.sp$num.rc.rs <- NA
for(i in 1:dim(aceitillar.sp)){
  
  aceitillar.sp$num.rc.rs[i] <-length(which(aceitillar.sp[i,j.cols] > 0))
  
}

hist(aceitillar.sp$num.rc.rs)


```


```{r}
aceitillar.sp$recapYN <- ifelse(aceitillar.sp$num.rc.rs > 0, "sp","not.sp")
  

```



## Look at ratio of caps to recaps
```{r}
with(aceitillar.sp, table(spp.code, recapYN) )
```

## 



```{r}
summary(factor(aceitillar.sp$stat))

aceitillar.sp$stat  <- as.character(aceitillar.sp$stat )
aceitillar.sp$stat.MvsR <- ifelse(aceitillar.sp$stat == "RE","R",aceitillar.sp$stat)

```



```{r}
i.1st.cap.96.N <- which(aceitillar.sp$N.96 == -1)
i.1st.cap.96.all <- which(aceitillar.sp$N.96 == -1 |
                          aceitillar.sp$J.97 == -1 |
                          aceitillar.sp$M.97 == -1)
length(i.1st.cap.96.N)
length(i.1st.cap.96.all)
# 
# summary(factor(aceitillar.sp[i.1st.cap.96.all,"yr.mo.1st.cap"]))
# summary(factor(aceitillar.sp[i.1st.cap.96.all,"recapYN"]))
# summary(factor(aceitillar.sp[i.1st.cap.96.all,"spp.code"]))
# summary(factor(aceitillar.sp[i.1st.cap.96.all,"spp.code"]))

m.sp.aceitillar <- bglmer(recapYN ~ 1 + stat.MvsR  +
                        (1|spp.code) +
                        (1|yr.mo.1st.cap), 
                   family = binomial,
                   data = aceitillar.sp[i.1st.cap.96.all,])
summary(m.sp.aceitillar)

m.sp.mencia.means <- bglmer(recapYN ~ -1 + stat.MvsR + 
                        (1|spp.code), 
                   family = binomial,
                   data = aceitillar.sp[i.1st.cap.96.all,])
summary(m.sp.mencia.means)
```




