---
title: "Regression Models of Condition (M.hat) Pasture sites"
author: "Nathan Brouwer"
date: "February 13, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries

```{r}
library(here)
fi <- here::here("R","load_libraries.R")
source(fi)
load_libraries()
```

## Load data

NOTE: not all species occur in Aceitillar!


```{r}
load(file = "./data/condition_Mhat.RData")

any(ls(pattern = "condition.Mhat") == "condition.Mhat")
```

### Set up predictor variable: site age

Check balanced of data
```{r}
with(condition.Mhat,
     table(year.cent, site.age))
```


```{r}
condition.Mhat$site.age.num <- condition.Mhat$site.age

#set Aceitillar to NA
condition.Mhat$site.age.num <- gsub("mature",NA,condition.Mhat$site.age.num)

condition.Mhat$site.age.num <- as.numeric(condition.Mhat$site.age.num)


condition.Mhat$site.age.cent <- with(condition.Mhat, 
                                 scale(site.age.num,
                                            scale = F))
```


### Code sites either Pasture or mature

THis was not used in final analysis

```{r}
#condition.Mhat$site.Y.G <- ifelse(condition.Mhat$site == "Aceitillar","mature","pasture")
```


## Model

Run  model w/ random slopes

#### Select focal spp

This removes rare spp
```{r}
i.mig <- which(condition.Mhat$stat.focals == "mig" &
              is.na(condition.Mhat$sex2) == FALSE)
i.res <- which(condition.Mhat$stat.focals == "res" &
              is.na(condition.Mhat$sex2) == FALSE)

i.all <- c(i.mig,i.res)
```

Check balance
```{r}
with(condition.Mhat[i.all,],
     table(spp.code, sex2))
```


Update index to Remove bad entries (very rare)
```{r}
i.BANA.bad <- with(condition.Mhat[,],
     which(spp.code == "BANA" & sex2 == "M"))
i.BAWW.bad <- with(condition.Mhat[,],
     which(spp.code == "BAWW" & sex2 == "Unknown"))

i.all <- i.all[-which(i.all %in% c(i.BANA.bad,i.BAWW.bad)) ]

```





### Regression Model for BLUPs using blmer()


#### Run model
```{r}
m.M.hat.0 <- blmer(M.hat.i ~ 1 + 
               (1|site) +
               (1|year) +
               (1|spp.code) +
               (site.age.cent|spp.code:sex2) , #+
               #(1|site:spp.code) +
               #(1|spp.code:site.Y.G), 
               control = lmerControl(optimizer = "Nelder_Mead"),
               data = condition.Mhat[i.all,])

summary(m.M.hat.0)
```

### Set up for plotting

 Extract BLUPs and their SEs
```{r}
m.hat.out <- data.frame(
 beta = ranef(m.M.hat.0)$"spp.code:sex2"[,2]
,se = se.ranef(m.M.hat.0)$"spp.code:sex2"[,2])

```


Approximate CIs
```{r}
m.hat.out$ci.hi <- m.hat.out$beta + 1.96*m.hat.out$se
m.hat.out$ci.lo <- m.hat.out$beta - 1.96*m.hat.out$se

```


Code approx "significance"
```{r}
m.hat.out$sig <- ifelse(m.hat.out[,"beta"] > 0 & m.hat.out[,"ci.lo"] > 0 |
       m.hat.out[,"beta"] < 0 & m.hat.out[,"ci.hi"] < 0, "*","" )
```


These are species w/ sufficent samples sizes / balance
```{r}
# focal migrants
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

# focal residents
focal.res <- c("HLCU","HILC",         #fix - which name to use?
               "STOF",
               "RLTH","NOMO",
               "GRWA",       "GTGT",   #fix - which name to use?
               "BANA","BCPT",
               "YFGR","BFGR",
               "GABU",
               "BWVI", #32; Aceitillar / Cueva +misc
               "GREP") #18 just Aceitillar - ever yr

focals <- c(focal.mig,
            focal.res)
```


Clean up df
```{r}
library(stringr)
ids <- str_split(rownames(m.hat.out),":",simplify = TRUE)
Species <- ids[,1]
Sex <- ids[,2]
Sex <- gsub("F","Female",Sex)
Sex <- gsub("M","Male",Sex)
rownames(m.hat.out) <- NULL
m.hat.out <- cbind(Species, Sex, m.hat.out)
```


Code residency
```{r}
m.hat.out$Status <- "Resident"
m.hat.out$Status[which(m.hat.out$Species %in% focal.mig)] <- "Migrant"
```



### Plot trends

To Do:
-add endemic status
-plot against condition
-calculate pasture age relative to year captured
-sort by magnitude
-add info on CIs

```{r}
pd <- position_dodge(0.75)
ggplot(dat = m.hat.out,
       aes(y = (beta),
           x = Species,
           color = Sex,
           shape = Sex)) +
  geom_errorbar(position= pd,
                aes(ymax = (beta+1*se),
                    ymin = (beta-1*se)),
                width = 0,
                size =2) +
  geom_errorbar(position= pd,
                aes(ymax = (ci.hi),
                    ymin = (ci.lo)),
                width = 0) +
      geom_point(position= pd,size = 4) +
  geom_hline(yintercept = (0)) +
  coord_flip() +
  facet_wrap(~Status,scales = "free") +
  ylab("Trend") +
  theme(legend.position = "bottom", 
    legend.direction = "horizontal",
    legend.justification =  "center",
    legend.title = element_blank())
```



### Model for overall effect
```{r}
m.M.hat.stat.focals <- update(m.M.hat.0, . ~ . + stat.focals )
m.M.hat.ADD         <- update(m.M.hat.0, . ~ . + stat.focals+site.age.cent)
m.M.hat.FULL        <- update(m.M.hat.0, . ~ . + stat.focals*site.age.cent)
```


```{r}
AICtab(m.M.hat.0,
       m.M.hat.stat.focals,
       m.M.hat.ADD,
       m.M.hat.FULL)
```


### Save analyses

```{r}
save.image("WRKSPC_condition_regression.RData")
```





