---
title: "Analysis: sex & age ratio - regression"
author: "Nathan Brouwer"
date: "January 17, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
dcast(data = ann_caps2,
      formula = spp.code ~ age.AHY.01,
      fun.aggregate = length,
      value.var = "age.AHY.01")
```



## Random slopes model - age ratio


### Model

Full random slopes model

```{r m.age.rand.slopes}
m.age.rand.slopes <- bglmer(as.numeric(age.AHY.01)-1 ~ 1 +
               (1|band) + #ann_caps2 contains all obs of each indivial
               (1|year) +
               (1|site) +
               #(1|site:spp.code) #this would be ideal but prob overkill
               (site.age.cent|spp.code),
              data = ann_caps2[-i.ann_caps2_aceit,],
              family = binomial#,
              #glmerControl(optimizer = "Nelder_Mead")
              )

summary(m.age.rand.slopes)
```


### Extract blus and calculate SEs
```{r}
out <- data.frame(beta = ranef(m.age.rand.slopes)[["spp.code"]]$site.age.cent,
se  = se.coef(m.age.rand.slopes)["spp.code"][["spp.code"]][,"site.age.cent"])

row.names(out) <- gsub("[ ][LE][al]","",row.names(out))
ids <- stringi::stri_split(row.names(out),regex= " ", simplify = T)

out$Species <- ids[,1]
out$Site <- ids[,1]
```


### Plot random slopes

From model m.age.rand.slopes
```{r}
out$Species <- factor(out$Species,
                      levels = focals)
out$Status <- "Resident"
out$Status[which(out$Species %in% focal.mig)] <- "Migrant"

ggplot(dat = out,
       aes(y = exp(beta),
           x = Species,
           color = Status)) +
  geom_point() +
  geom_errorbar(aes(ymax = exp(beta+1.96*se),
                    ymin = exp(beta-1.96*se)),
                width = 0) +
  geom_hline(yintercept = exp(0)) +
  coord_flip() +
  facet_wrap(~Status,scales = "free") +
  ylab("Trend") +
  ylim(0.90,1.075)
  
```







## Model overal trend by status (mig vs. resident)

### post-pasture sites

Aceitillar models seperatealy b/c it does not have a value for "site.age"

#### Null model

Similar to above but required tweaking to converged.  Current version should be identical but w/ different optimizer

```{r}
m.age.fixef.0 <- bglmer(as.numeric(age.AHY.01)-1 ~ 1 +
               (1|band) +
               (1|year) +
               (1|site) +
               #(1|site:spp.code) #this would be ideal but prob overkill
               (site.age.cent|spp.code),
              data = ann_caps2[-i.ann_caps2_aceit,],
              family = binomial,
              glmerControl(optimizer = "Nelder_Mead")
              )
```

#### Model fixed effects
```{r}
#status fixed effect
## (almost?) always converged
m.age.fixef.stat     <- update(m.age.fixef.0, . ~ . + status.focals)

#site.age.cent fixed effect
## throws warning
m.age.fixef.site.age <- update(m.age.fixef.0, . ~ . +           site.age.cent)

#convergence issues
##fails: with bobyqa under most/all conditions
##works: w/ Nelder_Mean when only intial captures are used AND when all recaps used
### even with Nelder_Mean does throw warning
m.age.fixef.add <- update(m.age.fixef.0, . ~ . + status.focals + site.age.cent)

#convergence issues
##fails with bobyqa under ALL conditions
##fails ...
##works: w/ Nelder_Mean when only intial captures are used 
##       
m.age.fixef.full <- update(m.age.fixef.0, . ~ . + status.focals*site.age.cent)

```


### Inference

```{r}
ICtab(m.age.fixef.0,
       m.age.fixef.stat,
       m.age.fixef.site.age, #conv warn
       m.age.fixef.add,      #fails
       m.age.fixef.full,
       base = TRUE,
       logLik = TRUE,
       type = "AICc")
```


### Plotting

#### Calculate CIs

```{r}
#merTools::predictInterval()

newdat <- expand.grid(
          band = m.age.fixef.full@frame$band[1]
          ,year = unique(m.age.fixef.full@frame$year)[1]
          ,site = unique(m.age.fixef.full@frame$site)[1]
          ,site.age.cent = unique(m.age.fixef.full@frame$site.age.cent)
          ,spp.code = unique(m.age.fixef.full@frame$spp.code)[1]
          ,status.focals = unique(m.age.fixef.full@frame$status.focals)
          )

names(m.age.fixef.full@frame)
out <- predictInterval(m.age.fixef.full, 
                       newdata =   newdat,
                which = "fixed", 
                level = 0.95,
                #n.sims = 10,
                stat = "median",
                include.resid.var = FALSE)

out <- cbind(newdat,out)

```


Determine mean of site.age variable to de-center it while plotting

```{r}
covar.mean <- mean(ann_caps2$site.age.init,na.rm = TRUE)
```

Plots
```{r}
gg.mencia <- ggplot(data = out,
        aes(y = invlogit(fit),
            x = site.age.cent+7.557782,
            color = status.focals)) +
  geom_line(aes(linetype = status.focals),
            size = 1) +
  xlab("Site age") +
  ylab("Age ratio (AHY vs all ages)") +
  geom_ribbon(aes(ymax = invlogit(upr),
                  ymin = invlogit(lwr),
                  fill = status.focals),
              alpha = 0.125,
              linetype = 0) +
  ylim(0,1) + theme(legend.position="none")
```




```{r}

```





