---
title: "Clean Aceitillar data for Mencia pasture succession project"
author: "Nathan Brouwer"
date: "January 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The Mencia pasture succession project uses part of the Aceitillar data as a reference.

This file loads, cleans, and reshapes the data


## Libraries
```{r}
library(here)
library(reshape2)
library(car)
library(stringi)
library(stringr)
library(lubridate)
```



## Load data
```{r}
skip <- 11
filename <- "Aceitillar_96-02_for_Mencia_all_caps_CSV.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  "CSV_Aceitillar_96-02",
                  filename)

aceit <- read.csv(filename, skip = 11,stringsAsFactors = F)
```



## Basic cleaning

### Change all headings to lower case

```{r}
names(aceit) <- tolower(names(aceit))
```

### Fix age

```{r}
aceit$age[-grep("AHY|HY",aceit$age)] <- NA
```


```{r}
summary(factor(aceit$age))
```

### Fix sex

```{r}
aceit$sex[-grep("F|M",aceit$sex)] <- NA
```


```{r}
#summary(factor(aceit$sex))
```


```{r}
#summary(factor(aceit$species))
```


### Fix typos in spp
```{r}
#remove space
aceit$species <- gsub("BCPT ", "BCPT", aceit$species)
```


```{r}
aceit$species[grep("Z",aceit$species)]

aceit$species <- gsub("ZEDO", "ZEND", aceit$species)
```

```{r}
aceit$species[grep("CGDO",aceit$species)]
```


```{r}
aceit$species <- gsub("CGDO", "COGD", aceit$species)
aceit$species <- gsub("CGOO", "COGD", aceit$species)
```


### HIPE Hisp Pewee
GAPE GREP
```{r}
aceit$species[grep("GREP",aceit$species)] <- "HIPE"
aceit$species[grep("GAPE",aceit$species)] <- "HIPE"
```


### Convert to factors

```{r}
aceit$colors <- factor(aceit$colors)
aceit$age <- factor(aceit$age)
aceit$sex <- factor(aceit$sex)

```


### Dates

Fix typo in name
```{r}
names(aceit)[grep("data.cap.orig", names(aceit))] <- "date.cap.orig"
```

Chang _ to -
```{r}
aceit$date.cap.raw <- aceit$date.cap.orig
aceit$date.cap.orig <- gsub("_","-",aceit$date.cap.orig)

```


Convert to standard R date format
```{r}
aceit$date.cap.orig <- mdy(aceit$date.cap.orig)
```


Compare dates
```{r}
#aceit[,c("date.cap.orig","date.cap.raw")]
```


### Order by date

Order by date 1st captured
```{r}
aceit <- aceit[order(aceit$date.cap.orig), ]
```


## Code focal migrants

These are the species focused on in the original reject MS

```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")
i <- which(aceit$species %in% focal.mig)

aceit$status.focals <- NA
aceit[i, "status.focals"] <- "mig"
```

### Focal residents

These are the species focused on in the original reject MS

```{r}
focal.res <- c("HLCU" #changes to HILC
,"STOF"
,"RLTH"
,"NOMO"
,"GRWA" # changes to GTGT
,"BANA"
,"BCPT"
,"YFGR"
,"BFGR"
,"GABU")
i <- which(aceit$species %in% focal.res)


aceit[i, "status.focals"] <- "res"
```

## Load spp meta / trait info

```{r}
spp.meta <- read.csv(here::here("data","spp_list.csv"),stringsAsFactors =F)
```


Subset columns I want to use
```{r}
spp.meta <- spp.meta[c("spp.code","spp",
                       "status2",
                       "hab1","diet")]

names(spp.meta) <- gsub("status2","status",names(spp.meta))
```





## Change GRWA -> GTGT in main dataframe

Green tailed ground warlber and green tailed ground tanager are the same

Change column name

```{r}
names(aceit) <- gsub("species","spp.code", names(aceit))
```


```{r}
aceit$spp.code <- as.character(aceit$spp.code)
aceit$spp.code[which(aceit$spp.code == "GRWA")] <- "GTGT"
```

```{r}
aceit$spp.code[which(aceit$spp.code == "HLCU")] <- "HILC"

```


MYWA =  myrtyle warbler = YRWA yellow rumped warbler
NUMA -> SBMU #nutmet manakin = scaly breasted munia
WHQD = WFQD

```{r}
aceit$spp.code[which(aceit$spp.code == "MYWA")] <- "YRWA"
aceit$spp.code[which(aceit$spp.code == "NUMA")] <- "SBMU"
aceit$spp.code[which(aceit$spp.code == "WHQD")] <- "WFQD"
```


## Merge cleaned data w/ species meta data


Merge
```{r}

#names(aceit)
#names(spp.meta)

aceit2 <- merge(aceit, spp.meta, all = T)

#dim(aceit)
#dim(aceit2)

aceit2 <-aceit2[-which(is.na(aceit2$site) == TRUE), ]
aceit <- aceit2
```


### Name site

Change current "site" column to "sub.site"

```{r}
aceit$sub.site <- aceit$site 
aceit$site <- "Aceitillar"
```

### Shorten column names

```{r}
names(aceit) <- gsub("nov","N",names(aceit))
names(aceit) <- gsub("jan","J",names(aceit))
names(aceit) <- gsub("feb","F",names(aceit))
names(aceit) <- gsub("mar","M",names(aceit))
names(aceit) <- gsub("wanderer","wndr",names(aceit))
names(aceit) <- gsub("species","spp",names(aceit))
names(aceit) <- gsub("year","yr",names(aceit))
names(aceit) <- gsub("month","mo",names(aceit))
names(aceit) <- gsub("band","bnd",names(aceit))
names(aceit) <- gsub("sex","sx",names(aceit))
names(aceit) <- gsub("colors","col",names(aceit))
names(aceit) <- gsub("notes","nts",names(aceit))
names(aceit) <- gsub("status","stat",names(aceit))
names(aceit) <- gsub("age","ag",names(aceit))
names(aceit) <- gsub("comments","nts",names(aceit))

```




## Code site persistence

```{r}
j <- grep("^[NDJFM]",names(aceit))
#names(aceit)[j]

x <- with(aceit,paste(N.96,J.97,M.97,N.97,J.98,M.98,J.99,
                 J.00,J.01,J.02,J.03),sep ="")
x <- gsub(" ","",x)

i.C <- which(x == "C")
x[i.C] <- "NP"
x[-i.C] <- "P"

aceit$persist <- x
```


## Save cleaned data

### Save .csv
```{r}
filename <- "aceit_site_persist_clean.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)

write.csv(aceit, file = filename,row.names = F)
```


```{r}

```

