---
title: "Reshape Aceitillar site persistence data to annual capture format"
author: "Nathan Brouwer"
date: "January 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries
```{r}
library(here)
library(reshape2)
library(car)
library(stringi)
library(lubridate)
```


## Load clean data


```{r}
filename <- "aceit_site_persist_clean.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)

aceit <- read.csv(file = filename,stringsAsFactors = F)
```



## Melt

Melt wide data to long format 
```{r}
names(aceit)
aceit.melt <- reshape2::melt(data = aceit,
     id.vars = c("spp.code",
                 "bnd","col","ag","sx",
                 "site",                   
                 "date.cap.orig",
                 "date.cap.raw",
                 "stat.focals",
                 "spp",
                 "stat",
                 "hab1","sub.site"    ,
                  "diet","persist",
                 "nts"), #formerly comments
     value.name = "cap.type")

```


```{r}
summary(factor(aceit.melt$cap.type))
summary(factor(aceit.melt$variable))
```


## Clean


### Remove blank entries

Blank entries for "cap.type" occur when a bird was not seen either b/c it had never been seen yet (before 1st cap) or after it had died/emmigrated/etc (after the last time it was seen)

```{r}
aceit.melt$cap.type  <- factor(ifelse(aceit.melt$cap.type == "",NA, aceit.melt$cap.type))


aceit.melt <- aceit.melt[-which(is.na(aceit.melt$cap.type == TRUE)),]

summary(factor(aceit.melt$cap.type))
```


### Extract date

Change "variable" column to date.cap for date captured
```{r}
names(aceit.melt)[which(names(aceit.melt) == "variable")] <- "date.cap"
```

Extract year

```{r}
temp <- stringr::str_split(aceit.melt$date.cap,"\\.",simplify = TRUE)

aceit.melt$month.cap <- temp[,1]
aceit.melt$year.cap <- temp[,2]
```

"year.cap"" is the year **actually** captured.
For analysis I need to determine the winter of sampling, which spans 2 calendar years, not the actualy year.  

So Nov  "1997" should become "1997-1998" b/c sampling that season continue into the next calendar year.

BUT Jan and March 1997 should be coded 1996-1997 b/c sampling began the calendar year prior. (coding this as 1997-1998 would put it into the wrong winter of sampling)

Note that for aceitillar sampling only ever occured in Nov, Jan and March, and that this only occured in 1996-1997, and 1997-1998.  In 1999 and beyond all sampling occurred in January.  Note that sampling in January 1999 is coded as "1998-1999" for consistency witht he prior years and the Mencia data.

```{r}
the90s <- as.character(c(96,97,98,99 ))

i.90s <- which(aceit.melt$year.cap %in% the90s) 

aceit.melt$year.cap[i.90s] <- paste("19",aceit.melt$year.cap[i.90s], sep = "")
aceit.melt$year.cap[-i.90s] <- paste("20",aceit.melt$year.cap[-i.90s], sep = "")


```


### Code winter of sampling
```{r}
summary(factor(aceit.melt$month))
```


```{r}
aceit.melt$year <- NA

#1996-1997
aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "1996" & 
                              month.cap == "N"))] <- "1996-1997"

aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "1997" &
                             month.cap %in% c("J","M")))] <- "1996-1997"

#1997-1998
aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "1997" & 
                              month.cap == "N"))] <- "1997-1998"

aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "1998" &
                             month.cap %in% c("J","M")))] <- "1997-1998"




## others
aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "1999" &
                             month.cap %in% c("J","M")))] <- "1998-1999"


aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "2000" &
                             month.cap %in% c("J","M")))] <- "1999-2000"

aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "2001" &
                             month.cap %in% c("J","M")))] <- "2000-2001"

aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "2002" &
                             month.cap %in% c("J","M")))] <- "2001-2002"

aceit.melt$year[with(aceit.melt, 
                     which(year.cap == "2003" &
                             month.cap %in% c("J","M")))] <- "2002-2003"

summary(factor(aceit.melt$year))

aceit.melt[is.na(aceit.melt$year),]
```

### Designate nominal year of capture

The 1st calendar year of the winter is the nominal year of that winter for the purpose of regression analysis

```{r}
aceit.melt$year.num <- gsub("^([12][90][901][1234567890])(.*)","\\1",
                            aceit.melt$year)
aceit.melt$year.num <- as.numeric(aceit.melt$year.num)
```

### Create date captured column

Date actually captured (not date orig captured)

Day not give; assume 15th of each month

```{r}
aceit.melt$date.cap <- paste(aceit.melt$month.cap,15,
          aceit.melt$year.cap,
          sep = "-")
aceit.melt$date.cap  <- gsub("J","Jan",aceit.melt$date.cap )
aceit.melt$date.cap  <- gsub("N","Nov",aceit.melt$date.cap )
aceit.melt$date.cap  <- gsub("M","Mar",aceit.melt$date.cap )

aceit.melt$date.cap  <- mdy(aceit.melt$date.cap )
```


### Order by date captured

```{r}
aceit.melt <- aceit.melt[order(aceit.melt$date.cap), ]
```

Check 
```{r}
aceit.melt[which(aceit.melt$bnd == "382"), ]
```


## Create band-year combo
This is down downstream when merged with Mencia

```{r}
# aceit.melt$ID.yr <- with(aceit.melt,
#                          paste(bnd,year,sep = "_")
#                          )
```



## Save ALL cleaned data

includes resight!

### Save .csv
```{r}
filename <- "aceit_annual_inc_resight.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)

write.csv(aceit.melt, file = filename,row.names = F)
```


## Remove resights

* captures are coded "C".
* resights are coded "S"
* if capture AND sighted: "CS"
* NOTE: the 1st capture of a year is NOT noted; a bird can be "C" in January and "C" again in Feb.

```{r}
dim(aceit.melt)

i.C <- grep("C",aceit.melt$cap.type)

summary(aceit.melt$cap.type[i.C])

aceit.melt.caps.only <- aceit.melt[i.C, ]

```



```{r}
max(summary(factor(aceit.melt.caps.only$bnd),10000))

aceit.melt.caps.only[grep("2_5900",aceit.melt.caps.only$bnd),]
```


## Save cleaned data

### Save .csv
```{r}
filename <- "aceit_annual.csv"
filename <- here::here("data-raw",
                  "data-raw-aceitillar_96-02",
                  filename)

write.csv(aceit.melt.caps.only, file = filename,row.names = F)
```


```{r}

```

