---
title: "Untitled"
author: "Nathan Brouwer"
date: "January 17, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Reconciling the two datasets

SL provided two complementary versions of the data

One was in seperate datasheets for each year

The other was organized to examine site persistance


At first glance the datasets appeared to have different numbers of unique records.  This is indeed true but is actualy expected.
Differences are almost all due to 2007-2008 records; this makes sense.  That was the last year of the study and so site persistance of newly captured birds cannot be determined


### Load cleaned and scrubbed LONG data

Data that was contained in seperate files by year
```{r}
filename <- "mencia_cleaned.csv"
filename <- here("data-raw",
                  "data-raw-mencia",
                  "mencia_all_captures_by_year",
                  filename)

mencia <- read.csv(filename)
```


### Load site persistence df
```{r}
filename <- "site_persist_cleaned.csv"
filename <- here("data-raw",
                  "data-raw-mencia",
                  "mencia_site_persistence",
                  filename)

sp <- read.csv(file = filename)
```




## Unique records in "mencia" dataframe
```{r}
length(unique(mencia$band))
```



## Unique records in "site persistance" dataframe

Each row should be unique.  There appear to be 3 duplicate bands.  This appears to be due to changes between sites within or between years.  

Note that the "menica" dataframe from annual records does contain at least some within-year recapture records, perhaps when there was movement between sites.


```{r}
i.unique.band <- match(unique(sp$band),sp$band)

sp$ID <- with(sp, paste(band,species,site, sep = "."))
i.unique.ID <- match(unique(sp$ID),sp$ID)

length(i.unique.band)
length(i.unique.ID)
dim(sp)
```


```{r}
sp$ID[duplicated(sp$ID)]

i.dup <- which(sp$ID %in% sp$ID[duplicated(sp$ID)])

sp[i.dup, 1:12]
```


      site band.pre band.suf colors species age  sex      date wing tarsus
1347 Cueva     2240    21542  XR-WR    AMRE  HY    F  4-Jan-04   57   17.5
1348 Cueva     2240    21542  XR-WR    AMRE  HY    F  5-Jan-04   NA     NA
1513 Cueva       4A      209  BX-YW    BCPT  HY <NA>  9-Nov-03   82   23.0
1517 Cueva       4A      209  BX-YW    BCPT  HY <NA>  6-Jan-04   83   23.9
1649 Cueva     2280    33228  WW-XR    CMWA AHY    F 18-Feb-04   64   18.4
1662 Cueva     2280    33228  WW-XR    CMWA AHY    F 22-Feb-04   NA     NA





### Look at these duplications in other dataframe

```{r}
dup.band <- sp$band[duplicated(sp$band)]

```

```{r}
mencia[which(mencia$band %in% dup.band), 1:12]
```


## Difference in number of unique records
```{r}
length(unique(mencia$band))-length(unique(sp$band))
```



## Are the species lists the same between years?

There are several spp not in the "site persistance" data
BWWA
LOST
SUTA
WHQD

```{r}
mencia.spp <- unique(mencia$species)
sp.spp <- unique(sp$species)

d1 <- data.frame(df = "mencia",
                 species = mencia.spp)
d2 <- data.frame(df = "sp",
                 species = sp.spp)

merge(d1,d2, by = "species",all = TRUE)
```


This only accounts for 7 records

```{r}
spp.x <- c("BWWA","LOST","SUTA","WHQD")

which(mencia$species %in% spp.x)

```



## Compare band lists

There appear to be 313 unique bands in the "mencia" dataframe that are not in the site persistance (sp) dataframe.  This matches the differences in dataframe size

```{r}
mencia.bands <- unique(mencia$band)
sp.bands <- unique(sp$band)

d1b <- data.frame(df = "mencia",
                 band = mencia.bands)
d2b <- data.frame(df = "sp",
                 band = sp.bands)

d3 <- merge(d1b,d2b, by = "band",all = TRUE)
```



Differences are almost all due to 2007-2008 records; this makes sense .  That was the last year of the study and so site persistance of newly captured birds cannot be determined

```{r}
bands.on.run <- d3[is.na(d3$df.y),"band"]

bands.on.run.i <- which(mencia$band %in% bands.on.run)

summary(mencia[bands.on.run.i, 1:19])
```


