---
title: "Ground insects from leaf litter"
author: "Nathan Brouwer"
date: "February 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

## Load Leaf litterl data

```{r}
load("./data/litter.RData")
library(reshape2)
library(vegan)
```



### Set factor levels

```{r}
litter$site <- factor(litter$site,
                           levels = c("La Cueva","La Caoba","Morelia","El Corral",
                                      "Broadleaf"))
```




## Reshape wide data to community matrix

```{r}
litter.comm.matrix <- dcast(data = litter,
                            formula = site + sample ~ taxa,
                            value.var = "N",
                            fun.aggregate = sum)

head(litter.comm.matrix)

```



REmove any samples that were all zeros
```{r}
row.sums. <- rowSums(litter.comm.matrix[,-c(1:2)])
i.0 <- which(row.sums. == 0)

litter.comm.matrix <- litter.comm.matrix[-i.0,]
```


## Reshape wide to total abundance

Total number of arthropods in a sample, ignoring taxa
```{r}
litter.tot.N <- dcast(data = litter,
                            formula = site + sample ~ .,
                            value.var = "N",
                            fun.aggregate = sum)
names(litter.tot.N)[3] <- "tot.N"
head(litter.tot.N)



```


## Plot

### Plot total abundance

```{r}
#mean SE/CIs on the fly
ggplot(data = litter.tot.N,
       aes(y = tot.N,
           x = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```


### Plot abundance by taxa

```{r}
litter <- merge(litter, litter.tot.N)

#mean SE/CIs on the fly
pd <- position_dodge(0.5)
i.unk <- which(litter$taxa== "Other")
ggplot(data = litter[,],
       aes(y = N,
           x = site,
           color = taxa,
           group = taxa#,
           #shape = taxa
           )) +
  stat_summary(position = pd,
               fun.data  = "mean_cl_boot",size = 2) +
  facet_wrap(~ taxa, scale = "free") +
  ggtitle("Leaf litter")
```



## Analyis


### Community analysis 

#### NMDS

Run NMDS
```{r}
litter_NMDS <- metaMDS(litter.comm.matrix[,-c(1:2)],
                        k=3,
                        maxit = 500,
                        try = 20, 
                        trymax = 50)

stressplot(litter_NMDS)
```


```{r}
ordiplot(litter_NMDS,xlim = c(-1,1))
ordiellipse(litter_NMDS,
         groups=litter.comm.matrix$site,
         label=T,
         kind = "se")

# ordibar(example_NMDS,
#          groups=litter.comm.matrix$site[-i.0],
#          kind = "se")
# 
# ordicluster(example_NMDS,
#          groups=litter.comm.matrix$site[-i.0],
#          label=T,
#          kind = "se")
```






### Model total abundance
```{r}
m.tot.N.0 <- glmer(tot.N ~ 1 + (1|site), data = total.N, family = "poisson")
m.tot.N.site <- glmer(tot.N ~ site + (1|site), data = total.N, family = "poisson")
m.tot.N.means <- glmer(tot.N ~ -1 + site + (1|site), data = total.N, family = "poisson")
anova(m.tot.N.0,
      m.tot.N.site)

coefplot(m.tot.N.means,
         xlab = "Relative Arthropod Abundance (log(N)-ish)",
         horizontal = TRUE,
         ylab = "Site",
         title = "Relative Arthropod Abundance (log(N)-ish)")
```



```{r}
ggplot(data  = x,
       aes(y = fit,
           x = site)) +
  geom_point() +
  geom_errorbar(aes(ymax = upr,
                    ymin = lwr),
              width = 0)
```

```{r}
coef(m.simps.0)
ranef(m.simps.site)
fixef(m.simps.means)
```




## MANOVA

### MANOVA - raw counts
```{r}
manova.0 <- manova(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ 1, 
                  data = litter.comm.matrix)

manova.site <- manova(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ site, 
                  data = litter.comm.matrix)
anova(manova.0,
      manova.site)
summary(manova.site, test="Pillai")


```

#### Univariate follow-up ANOVAs

Ant 0.016
Cockroach = 0.057
Gusano = 0.068 (worms)

```{r}
summary.aov(manova.site)
anova.ants <- lm(Ants ~ site , 
                  data = litter.comm.matrix)

aov.ants <- aov(Ants ~ site + Error(site), 
                  data = litter.comm.matrix)

```


## PERMANOVA - raw counts

"manhattan", "euclidean", "canberra", "bray", "kulczynski", "jaccard", "gower", "altGower", "morisita", "horn", "mountford", "raup" , "binomial", "chao", "cao" or "mahalanobis"

```{r}
row.sums. <- rowSums(litter.comm.matrix[,-c(1:2)])
i.0 <- which(row.sums. == 0)
adonis.site <- adonis(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ site,
                      groups = "site",
                  data = litter.comm.matrix[-i.0,])

adonis.site$aov.tab
adonis.site$coefficients
```


###

```{r}
man <- mvabund(litter.comm.matrix[,-c(1:2)]) 
plot(man ∼ litter.comm.matrix$site)

many. <- manyglm(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ site,
                      groups = "site",
                  data = litter.comm.matrix[,])

plot(many.)
plot(many. ∼ site)
```




## Percentages

```{r}
divideit <- function(x){x/row.tot}
row.tot <- rowSums(litter.comm.matrix[,-c(1:2)])

litter.comm.matrix.percent <- litter.comm.matrix
for(i in 1:length(row.tot)){
  litter.comm.matrix.percent[i,-c(1:2)] <- litter.comm.matrix.percent[i,-c(1:2)]/row.tot[i]
}

litter.comm.matrix.percent <- litter.comm.matrix.percent[-which(row.tot == 0),]

```




## MANOVA

### MANOVA - raw counts
```{r}
manova.percent.0 <- manova(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ 1, 
                  data = litter.comm.matrix.percent)

manova.percent.site <- manova(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ site, 
                  data = litter.comm.matrix.percent)
anova(manova.percent.0,
      manova.percent.site)
summary(manova.percent.site, test="Pillai")


```

#### Univariate follow-up ANOVAs

Ant 0.016
Cockroach = 0.057
Gusano = 0.068 (worms)

```{r}
summary.aov(manova.percent.site)
anova.ants <- lm(Ants ~ site, 
                  data = litter.comm.matrix)
```




## PERMANOVA - percentage

"manhattan", "euclidean", "canberra", "bray", "kulczynski", "jaccard", "gower", "altGower", "morisita", "horn", "mountford", "raup" , "binomial", "chao", "cao" or "mahalanobis"

```{r}

adonis.site.percent <- adonis(cbind(Ants,Beetle,Cockroach,
                        Cricket,Gusano,Other,Spiders) ~ site,
                      groups = "site",
                  data = litter.comm.matrix.percent)

adonis.site.percent$aov.tab
adonis.site$coefficients
```



### Plot perecentage

```{r}
Ants <- ggplot(data = litter.comm.matrix.percent,
       aes(y = Ants,
           x = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2) +
  # geom_boxplot() +
  # geom_point() +
  ggtitle("Pertange of ants in sample")

Spiders <- ggplot(data = litter.comm.matrix.percent,
       aes(y = Spiders,
           x = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2) +
  # geom_boxplot() +
  # geom_point() +
  ggtitle("Pertange of Spiders in sample")

Cockroach <- ggplot(data = litter.comm.matrix.percent,
       aes(y = Cockroach,
           x = site)) +
  # geom_boxplot() +
  # geom_point() +
  stat_summary(fun.data  = "mean_cl_boot",size = 2) +
  ggtitle("Pertange of Cockroach in sample")


Beetle <- ggplot(data = litter.comm.matrix.percent,
       aes(y = Beetle,
           x = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2) +
  # geom_boxplot() +
  # geom_point() +
  ggtitle("Pertange of Beetle in sample")

cowplot::plot_grid(Ants,Spiders,Cockroach,Beetle)
```


```{r}
which(litter.comm.matrix.percent$Ants == 0)
which(litter.comm.matrix.percent$Ants == 1)

(m1 <- glmmTMB(Ants~ site + (1|site),
family=betar, 
data=litter.comm.matrix.percent))
summary(m1)

```



### logit transform percenage
```{r}
library(car)
library(blme)
summary(lm(logit(Ants) ~ site , 
             data = litter.comm.matrix.percent))

summary(blmer(logit(Ants) ~ -1+ librsite + (1|site), 
             data = litter.comm.matrix.percent))
```

