---
title: "Ground arthropods from leaf litter"
author: "Nathan Brouwer"
date: "February 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

GLMM and MANOVA analyses of leaf litter insects.  

See seperate script for flying insects from sticky traps.

## Load Leaf litter data

```{r}
load("./data/litter.RData")
source('~/1_R/git/git-aviary/DRmencia/R/back_transform_reals.R')
library(reshape2)
library(vegan)
library(lme4)
library(blme)
library(ggplot2)
library(cowplot)
library(merTools)
```



## Reshape wide data to community matrix

Data is long; reshape to wide for MANOVA
```{r}
litter.comm.matrix <- dcast(data = litter,
                            formula = site + sample ~ taxa,
                            value.var = "N",
                            fun.aggregate = sum)

head(litter.comm.matrix)

litter.comm.matrix <- na.omit(litter.comm.matrix)

```




## Reshape wide to total abundance

Total number of arthropods in a sample, ignoring taxa

NOte: "Broadleaf" = Aceitillar
```{r}
litter.tot.N <- dcast(data = litter,
                            formula = site + sample ~ .,
                            value.var = "N",
                            fun.aggregate = sum)
names(litter.tot.N)[3] <- "tot.N"


### set factor levels
litter.tot.N$site <- factor(litter.tot.N$site,
                        levels =c("La Cueva",
                                  "La Caoba",
                                  "Morelia",
                                  "El Corral",
                                "Broadleaf"))
```


### Set ages for regression

Will use nested regression ; no age for Broadleaf site
```{r}
litter.tot.N$site.age.init <- NA
litter.tot.N$site.age.init[which(litter.tot.N$site == "La Cueva")] <- 2
litter.tot.N$site.age.init[which(litter.tot.N$site == "La Caoba")] <- 5
litter.tot.N$site.age.init[which(litter.tot.N$site == "Morelia")] <- 10 
litter.tot.N$site.age.init[which(litter.tot.N$site == "El Corral")] <- 20
```


## Plot

### Plot total abundance

#### Plot means

Exploratory plot
```{r}
#mean SE/CIs on the fly
ggplot(data = litter.tot.N,
       aes(y = tot.N,
           x = site)) +
  #ylim(25,85) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```


#### Plot means w/ age as continous
Exploratory plot.  This is how data will be modeled.
```{r}
pd <- position_dodge(0.5)

ggplot(data = litter.tot.N[,],
       aes(y = tot.N,
           x = site.age.init)) +
  #ylim(25,75) +
  geom_smooth(method=glm,
              method.args = list(family = "poisson"))+
  #geom_point(position = pd) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```


####  Plot percentages

```{r}
litter2 <- merge(litter,litter.tot.N, all =T)
```

Clean up to make easier to view axes
```{r}
litter2$site2 <- litter2$site
litter2$site2 <- gsub("La ","",litter2$site2)
litter2$site2 <- gsub("El ","",litter2$site2)
litter2$site2 <- gsub("Broadleaf","Aceitillar",litter2$site2)


litter2$site2 <- factor(litter2$site2,
                        levels =c("Cueva",
                                  "Caoba",
                                  "Morelia",
                                  "Corral",
                                "Aceitillar"))
```


Plot percentages for exploraiton
```{r}
ggplot(data = (litter2),
       aes(y = N/tot.N,
           x = site2,
           color = site2)) +
  stat_summary(fun.data  = "mean_cl_boot") +
  facet_wrap(~taxa,
             scale = "free") +
  xlab("Site") +
  ylab("Percentage composition") +
  ggtitle("Leaf litter arthropods")
    
```


## Model Litter arthropod data

### Set up
```{r}
#observation level random effect for overdispersion
litter.tot.N$i <- 1:dim(litter.tot.N)[1]
litter.tot.N$year <- litter.tot.N$site.age.init
i.aceit <- which(litter.tot.N$site == "Broadleaf")
```

### Run regression model

Reported in draft MS as "Across pasture sites litter arthropod abundance increased marginally with age (β=0.039, SE= 0.022, χ23,4 = 4.03, PLRT = 0.045)."
```{r}
#model null
litter.regress0 <- bglmer(tot.N ~ 1 +
                (1|site) +
                (1|i),
              family = poisson,
              data = litter.tot.N[-i.aceit,])

#add site age
litter.regress1 <- update(litter.regress0, 
                          . ~ . + site.age.init)

summary(litter.regress1)

anova(litter.regress0,
      litter.regress1)

```


## ANOVA

1-way ANOVA.  This is not used for inference, but for producing estimated means for plotting.

```{r}
litter.anova.bglmm <- bglmer(tot.N ~  site +
                #(1|site) +
                (1|i),
              family = poisson,
              data = litter.tot.N)

litter.anova.bglmm.means <- bglmer(tot.N ~  -1 + site +
                #(1|site) +
                (1|i),
              family = poisson,
              data = litter.tot.N)

```



## Poisson t-test-ish

Build models which pools all pasture sites and compares to forest.


Written up in draft MS as "Leaf litter insects were 0.55 (Bootstrapped 95% CI 0.22 -1.32) times as abundant in the mature forest compared to the pastures, though the difference was only marginally significant (χ23,4 = 2.69, PLRT =0.10).  "

### Model
```{r}
#make factor that designated Aceit vs. pasture
litter.tot.N$Aceit.YN <- ifelse(litter.tot.N$site == "Broadleaf",
                                "Aceit.YES",
                                "Aceit.NO")

#model null
litter.pasture.vs.broadleaf.0 <- bglmer(tot.N ~ 1 +
                                         (1|site) +
                (1|i),
              family = poisson,
              data = litter.tot.N[,])

#model effect
litter.pasture.vs.broadleaf.1 <- update(litter.pasture.vs.broadleaf.0,
                                        .~. + Aceit.YN)

#test 
anova(litter.pasture.vs.broadleaf.0,
      litter.pasture.vs.broadleaf.1)

```


### Bootstrap CIs
Calcualte CIs: determine by what factor they are different


This throws a few warnings for a few of the bootstrap samples but the max gradients are still small.  Re-running with Nelder_Mead should fix this.

Exact confidnece interval coverage is slightly unstable; Re-running w/ betetr optimizer and with more bootstraps should clear up.

```{r}
#ci.out.litter <-  confint(litter.pasture.vs.broadleaf.1,method = "boot")

#ci.out.litter
```
Output:

                       2.5 %    97.5 %
.sig01             0.8595836 1.1941171
.sig02             0.1290505 0.5626109
(Intercept)        1.1628157 1.8979039
Aceit.YNAceit.YES -1.5189240 0.2761946

Exponentiated:
0.2189473 - 0.55- 1.318104


## Plot regression model

Build up plot with regression output, means, CIs, and raw data.

For other analyses I built these using functions; here I adapted the code for those functions ad hoc.


### Data for prediction

#### Regression output
```{r}
newdat.reg.litter <- expand.grid(
     site.age.init = seq(min(litter.regress1@frame$site.age.init),
                       max(litter.regress1@frame$site.age.init))
    ,site = unique(litter.regress1@frame$site)[1]
    ,i = litter.regress1@frame$i[1]
    )



```

##### ANOVA
```{r}
newdat.anova.litter.bglmm <- expand.grid(
      site =unique(litter.anova.bglmm@frame$site)
      ,litter.anova.bglmm@frame$site[1]
      ,i = litter.anova.bglmm@frame$i[1]
    )
```




#### Predictions from model using merTools::predictInterval


# Regression model
```{r}
litter.rgrxn.pred <- merTools::predictInterval(litter.regress1,
                                   newdata =   newdat.reg.litter,
                                   which = "fixed",
                                   level = 0.95,
                                   #n.sims = 10,
                                   stat = "median",
                                   include.resid.var = FALSE)

litter.rgrxn.pred2 <- back_transform_reals(litter.rgrxn.pred,fnxn = "exp")
litter.rgrxn.pred3 <- cbind(newdat.reg.litter,litter.rgrxn.pred2)

```

# Anova model
```{r}
litter.anova.pred <- merTools::predictInterval(litter.anova.bglmm,
                                   newdata =   newdat.anova.litter.bglmm,
                                   which = "fixed",
                                   level = 0.95,
                                   #n.sims = 10,
                                   stat = "median",
                                   include.resid.var = FALSE)

litter.anova.pred2 <- back_transform_reals(litter.anova.pred,fnxn = "exp")

litter.anova.pred3 <- cbind(newdat.anova.litter.bglmm,litter.anova.pred2)

```


```{r}
summary(litter.anova.bglmm.means)
exp(1.0183+0.2221*1.96)
exp(1.0183-0.2221*1.96)

```



## Plot
```{r}
ymin. <- 0
ymax.<- 25

gg.litter.regress <-  ggplot(data = litter.rgrxn.pred3,
         aes(y = fit.real,
             x = site.age.init,
             color = site,
             shape = site)) +
    geom_line(aes()) +
    ylim(ymin.,ymax.) +
    xlab("Site age") +
    ylab("Litter arthropod abundance") +
    scale_shape_manual(values=c(4,0, 1, 2, 5)) +

    geom_ribbon(aes(ymax = upr.real,
                    ymin = lwr.real),
                alpha = 0.05,
                linetype = 0)  +
    #ylim(ymin., ymax.) +
    theme(legend.position="none")

## add raw data
gg.litter.regress1 <- gg.litter.regress + 
       geom_point(data = litter.regress1@frame,
                      aes(y = litter.regress1@frame[,1],
                          x = site.age.init),
                      size = 3)

## add ANOVA dat
nudge. <- 0.5
litter.anova.pred3$site.age.init <- NA
litter.anova.pred3$site.age.init[which(litter.anova.pred3$site == "La Cueva")] <- 2
litter.anova.pred3$site.age.init[which(litter.anova.pred3$site == "La Caoba")] <- 5
litter.anova.pred3$site.age.init[which(litter.anova.pred3$site == "Morelia")] <- 10 
litter.anova.pred3$site.age.init[which(litter.anova.pred3$site == "El Corral")] <- 20

gg.litter.regress2 <- gg.litter.regress1 + geom_point(data = litter.anova.pred3,
                                aes(y = fit.real,
                                    x =  site.age.init+nudge.),
                                size = 5,
                                shape =16,
                                position = pd) +
                 geom_errorbar(data = litter.anova.pred3,
                               aes(x = site.age.init+nudge.,
                                  ymax = upr.real,
                                   ymin = lwr.real),
                               width = 0)
```



```{r}
  pd <- position_dodge(0.25)

litter.anova.pred3$site <- gsub("Broadleaf",
                                "Aceitillar",
                                litter.anova.pred3$site)

litter.anova.bglmm@frame$site <- gsub("Broadleaf",
                                "Aceitillar",
                                litter.anova.bglmm@frame$site)


i.broad.lit <- which(litter.anova.pred3$site == "Aceitillar")
i.broad2.lit <- which(litter.anova.bglmm@frame$site == "Aceitillar")

y.mean <- litter.anova.pred3[i.broad.lit,"fit.real"]
ymin.CI <- litter.anova.pred3[i.broad.lit,"lwr.real"]
ymax.CI <- litter.anova.pred3[i.broad.lit,"upr.real"]


gg.anova.litter <-  ggplot(data = litter.anova.pred3[i.broad.lit,],
                    aes(y = fit.real,
                        x = site)) +
    geom_line(aes(),
              size = 1) +
    ylim(ymin.,ymax.) +
    xlab("") +
    ylab("") + 
  geom_jitter(data = litter.anova.bglmm@frame[i.broad2.lit,],
                                #position = pd,
                                width  = 0.125,
                                shape = 10,
                                size = 3,
                                aes(y = litter.anova.bglmm@frame[i.broad2.lit,1],
                                    x = site)) +
    theme(axis.line.y=element_blank(),
          #axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          #axis.title.x=element_blank(),
          axis.title.y=element_blank()) +
  
  annotate("pointrange", x = 1.25, 
           y = y.mean,
           ymin = ymin.CI, 
           ymax = ymax.CI,
  size = 1.5,
  fatten = 2)


# gg.anova.litter2 <- gg.anova.litter + geom_errorbar(position = pd,
#                   aes(ymax = upr.real,
#                     ymin = lwr.real),
#                 width = 0)  +
#     geom_point(position = pd,
#                size = 5,
#                shape = 1) +
#     theme(legend.position="none") +
#     theme(axis.line.y=element_blank(),
#           #axis.text.x=element_blank(),
#           axis.text.y=element_blank(),
#           axis.ticks=element_blank(),
#           #axis.title.x=element_blank(),
#           axis.title.y=element_blank())
```


```{r}
cowplot::plot_grid(gg.litter.regress2,
                   gg.anova.litter,
                   rel_widths = c(8,1))
```

### Save plots
```{r}
save(gg.litter.regress2, file = "gg_litter_insects_regression.RData")
save(gg.anova.litter, file = "gg_litter_insects_anova.RData")
```

## Analyis


### Manova

```{r}
with(na.omit(litter.comm.matrix), table(site, sample))
i.drop <- which(litter.comm.matrix$sample > 10)

manova.dat.litter <- litter.comm.matrix[-i.drop,]
manova.out.litter <- manova(cbind(Ants,Beetle,
                           Cockroach,Cricket,Gusano,
                           Other,Spiders) ~ site, 
                           data = manova.dat.litter )
summary(manova.out.litter)
summary.aov(manova.out.litter)
```

          Df Pillai approx F num Df den Df   Pr(>F)   
site       4 1.1174   1.9867     32    164 0.002933 **
Residuals 45                                          


## Save workspac

```{r}
save.image(file = "WORKSPACE_leaf_litter_insect_analysis.RData")
```

