---
title: "Flying insects from sticky traps"
author: "Nathan Brouwer"
date: "February 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

GLMM and MANOVA analyses of leaf litter insects.  See seperate script for flying insects from sticky traps.  Annotation in that script is generally more thorough than this one.


## Load Leaf flyingl data

```{r}
load("./data/flying.RData")
library(reshape2)
library(vegan)
library(lme4)
library(blme)
library(ggplot2)
library(cowplot)
library(merTools)
source('~/1_R/git/git-aviary/DRmencia/R/back_transform_reals.R')
```


## Reshape wide data to community matrix

```{r}
flying.comm.matrix <- dcast(data = flying,
                            formula = site + sample ~ taxa,
                            value.var = "N",
                            fun.aggregate = sum)

#remove any rows that are all zero; 
flying.comm.matrix <- na.omit(flying.comm.matrix)

```

## Reshape wide to total abundance

Total number of arthropods in a sample, ignoring taxa
```{r}

flying.tot.N <- dcast(data = flying,
                            formula = site + sample ~ .,
                            value.var = "N",
                            fun.aggregate = sum)
names(flying.tot.N)[3] <- "tot.N"


#reset factor levels
flying.tot.N$site2 <- factor(flying.tot.N$site,
                        levels =c("La Cueva",
                                  "La Caoba",
                                  "Morelia",
                                  "El Corral",
                                "Aceitillar"))
```



### Set ages for regression

Not age for the broadleaf forest at Aceitillar
```{r}
flying.tot.N$site.age.init <- NA
flying.tot.N$site.age.init[which(flying.tot.N$site == "La Cueva")] <- 2
flying.tot.N$site.age.init[which(flying.tot.N$site == "La Caoba")] <- 5
flying.tot.N$site.age.init[which(flying.tot.N$site == "Morelia")] <- 10 
flying.tot.N$site.age.init[which(flying.tot.N$site == "El Corral")] <- 20
```


## Plot

### Plot total abundance


#### Plot means
```{r}
#mean SE/CIs on the fly
ggplot(data = flying.tot.N,
       aes(y = tot.N,
           x = site)) +
  #ylim(25,85) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```


## Plot means w/ age as continous
```{r}
pd <- position_dodge(0.5)
i.max <- which(flying.tot.N$tot.N > 100 |
               flying.tot.N$site == "Broadleaf" |
               is.na(flying.tot.N$tot.N) == TRUE)
ggplot(data = flying.tot.N[-i.max,],
       aes(y = tot.N,
           x = site.age.init)) +
  ylim(25,75) +
  geom_smooth(method=lm)+
  #geom_point(position = pd) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```


##  Calculate percentages

```{r}
flying2 <- merge(flying,flying.tot.N)

dim(flying2)

dim(flying)
```


Plot percentages
```{r}

flying2$site2 <- flying2$site
flying2$site2 <- gsub("La ","",flying2$site2)
flying2$site2 <- gsub("El ","",flying2$site2)
flying2$site2 <- gsub("Broadleaf","Aceitillar",flying2$site2)

summary(factor(flying2$site))
summary(factor(flying2$site2))
flying2$site2 <- factor(flying2$site2,
                        levels =c("Cueva",
                                  "Caoba",
                                  "Morelia",
                                  "Corral",
                                "Aceitillar"))
summary(flying$site)

ggplot(data = flying2,
       aes(y = N/tot.N,
           x = site2)) +
  stat_summary(fun.data  = "mean_cl_boot") +
  facet_wrap(~taxa,
             scale = "free") +
  xlab("Site") +
  ylab("Percentage composition") +
  ggtitle("Flying arthropods")
    
```


## Model

```{r}
flying.tot.N$i <- 1:dim(flying.tot.N)[1]
flying.tot.N$year <- flying.tot.N$site.age.init
i.aceit <- which(flying.tot.N$site == "Broadleaf")
flying.regress0 <- bglmer(tot.N ~ 1 +
                (1|site) +
                (1|i),
              family = poisson,
              data = flying.tot.N[-i.aceit,])

flying.regress1 <- update(flying.regress0, 
                          . ~ . + site.age.init)

anova(flying.regress0,
      flying.regress1)

i.max2 <- which(flying.tot.N$tot.N > 100 |
               #flying.tot.N$site == "Broadleaf" |
               is.na(flying.tot.N$tot.N) == TRUE)

#don't use means parameterization; doesn't play nice
#with downstream plotting
flying.anova <- bglmer(tot.N ~  site +
                #(1|site) +
                (1|i),
              family = poisson,
              data = flying.tot.N[,])
```




BUild models which pool all pasture sites

```{r}
#make factor that designates Aceit vs. pasture
flying.tot.N$Aceit.YN <- ifelse(flying.tot.N$site == "Broadleaf","Aceit.YES",
                                "Aceit.NO")

#model
flying.pasture.vs.broadleaf.0 <- bglmer(tot.N ~ 1 +
                                         (1|site) +
                (1|i),
              family = poisson,
              data = flying.tot.N[-i.max2,])
flying.pasture.vs.broadleaf.1 <- update(flying.pasture.vs.broadleaf.0,
                                        .~. + Aceit.YN)

#test 
anova(flying.pasture.vs.broadleaf.0,
      flying.pasture.vs.broadleaf.1)




```


Calcualte CIs: determine by what factor they are different
```{r}
#ci.out <-  confint(flying.pasture.vs.broadleaf.1,method = "boot")
```
                       2.5 %    97.5 %
.sig01            0.19080235 0.3162773
.sig02            0.06657935 0.3454228
(Intercept)       3.43297028 3.8569679
Aceit.YNAceit.YES 0.02347341 0.9440533







### Plot

```{r}
# data for prediction
newdat.flying.reg <- expand.grid(
     site.age.init = seq(min(flying.regress1@frame$site.age.init),
                       max(flying.regress1@frame$site.age.init))
    ,site = unique(flying.regress1@frame$site)[1]
    ,i = flying.regress1@frame$i[1]
    )

newdat.flying.anova <- expand.grid(
      site = unique(flying.anova@frame$site)
      ,i = flying.anova@frame$i[1]
    )

```

```{r}
## predictions
flying.rgrxn.pred <- merTools::predictInterval(flying.regress1,
                                   newdata =   newdat.flying.reg,
                                   which = "fixed",
                                   level = 0.95,
                                   #n.sims = 10,
                                   stat = "median",
                                   include.resid.var = FALSE)

flying.rgrxn.pred2 <- back_transform_reals(flying.rgrxn.pred,fnxn = "exp")

flying.rgrxn.pred3 <- cbind(newdat.flying.reg
                            ,flying.rgrxn.pred2)

```


```{r}
flying.anova.pred <- merTools::predictInterval(flying.anova,
                                   newdata =   newdat.flying.anova,
                                   which = "fixed",
                                   level = 0.95,
                                   #n.sims = 10,
                                   stat = "median",
                                   include.resid.var = FALSE)

flying.anova.pred2 <- back_transform_reals(flying.anova.pred,fnxn = "exp")
flying.anova.pred3 <- cbind(newdat.flying.anova,flying.anova.pred2)

```



```{r}
ymin. <- 15
ymax.<- 100

gg.out.fly.reg <-  ggplot(data = flying.rgrxn.pred3,
         aes(y = fit.real,
             x = site.age.init,
             color = site,
             shape = site)) +
    geom_line(aes()) +
   ylim(ymin.,ymax.) +
    xlab("Site age") +
    ylab("Flying arthropod abundance") +
   scale_shape_manual(values=c(4,0, 1, 2, 5)) +

    geom_ribbon(aes(ymax = upr.real,
                    ymin = lwr.real),
                alpha = 0.05,
                linetype = 0)  +
    #ylim(ymin., ymax.) +
    theme(legend.position="none")

  ## add raw data
  gg.out.fly.reg1 <- gg.out.fly.reg + geom_point(data = flying.regress1@frame,
                      aes(y = flying.regress1@frame[,1],
                          x = site.age.init),
                      size = 3)

  ## add ANOVA dat
  nudge. <- 0.5

  flying.anova.pred3$site.age.init <- c(2,5,10,20, NA)
  gg.out.fly.reg2 <- gg.out.fly.reg1 + geom_point(data = flying.anova.pred3,
                                aes(y = fit.real,
                                    x =  site.age.init+nudge.),
                                size = 5,shape =16,
                                position = pd) +
                 geom_errorbar(data = flying.anova.pred3,
                               aes(x = site.age.init+nudge.,
                                  ymax = upr.real,
                                   ymin = lwr.real),
                               width = 0)
```



```{r}
flying.anova.pred3$site <- gsub("Broadleaf",
                                "Aceitillar",
                                flying.anova.pred3$site)

flying.anova@frame$site <- gsub("Broadleaf",
                                "Aceitillar",
                                flying.anova@frame$site)


i.broad <- which(flying.anova.pred3$site == "Aceitillar")
i.broad2 <- which(flying.anova@frame$site == "Aceitillar")

y.mean <- flying.anova.pred3[i.broad,"fit.real"]
ymin.CI <- flying.anova.pred3[i.broad,"lwr.real"]
ymax.CI <- flying.anova.pred3[i.broad,"upr.real"]

gg.anova.fly <-  ggplot(data = flying.anova.pred3[i.broad,],
                    aes(y = fit.real,
                        x = site)) +
    geom_line(aes(),
              size = 1) +
    ylim(ymin.,ymax.) +
     xlab("") +
    # ylab(ylab.) +
  annotate("pointrange", x = 1.25, 
           y = y.mean,
           ymin = ymin.CI, 
           ymax = ymax.CI,
  size = 1.5,
  fatten = 2) +
    # geom_errorbar(aes(ymax = upr.real,
    #                 ymin = lwr.real),
    #             width = 0)  +
    # geom_point(size = 3) +
    theme(legend.position="none")# +
    #ylim(ymin., ymax.)

  ## add raw data
#  pd <- position_dodge(0.25)
gg.anova.fly2 <- gg.anova.fly + geom_jitter(data = flying.anova@frame[i.broad2,],
                                #position = pd,
                                width  = 0.125,
                                shape = 10,
                                aes(y = flying.anova@frame[i.broad2,1],
                                    x = site)) +
    theme(axis.line.y=element_blank(),
          #axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          #axis.title.x=element_blank(),
          axis.title.y=element_blank())

```


### Check plots

```{r}
cowplot::plot_grid(gg.out.fly.reg2,
                   gg.anova.fly2,
                   rel_widths = c(8,1))
```


### Save plots
```{r}
save(gg.out.fly.reg2, file = "gg_flying_insects_regression.RData")
save(gg.anova.fly2, file = "gg_flying_insects_anova.RData")
```


## Analyis


### Manova

```{r}
with(na.omit(flying.comm.matrix), table(site, sample))
i.drop <- which(flying.comm.matrix$sample > 10)

manova.dat <- flying.comm.matrix[-i.drop,]
manova.out <- manova(cbind(Coleoptera,Diptera,Homoptera,Hymenoptera,
             Lepidoptera,Orthoptera,Spiders,
             Unidentified) ~ site, data = manova.dat )
summary(manova.out)
summary.aov(manova.out)
```

          Df Pillai approx F num Df den Df   Pr(>F)   
site       4 1.1174   1.9867     32    164 0.002933 **
Residuals 45                                          



## Save workspac

```{r}
save.image(file = "WORKSPACE_leaf_litter_insect_analysis.RData")
```

