---
title: "Make Tables 5 & 6: Site persistence"
author: "Nathan Brouwer"
date: "January 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script create the site persistence parts tables 5 and 6
It uses a dataset where Mencia and Aceitillar data have been combined.

## Libraries
```{r}
library(reshape2)
library(here)
```



## Load data

Load annual Captures data with Mencia AND Aceitillar combined


```{r}
filename <- "site_persist.RData"
load(here::here("data",filename))
ls()
```


## Subset by focal species


```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

focal.res <- c("HILC" # formerly HLCU
,"STOF"
,"RLTH"
,"NOMO"
,"GTGT"   # formerly GRWA
,"BANA"
,"BCPT"
,"YFGR"
,"BFGR"
,"GABU")

focals <- c(focal.mig,focal.res)
```

```{r}
i.focal <- which(site_persist$spp.code %in% focals)
site_persist.foc <- site_persist[i.focal,]
```


## Set order for tables

replicate order from  tables in oridingal MS

```{r}
site_persist.foc$spp.code <- factor(site_persist.foc$spp.code,
                             levels = focals)
```


## Order sites by age

```{r}
site_persist.foc$site <- factor(site_persist.foc$site,
                      levels = c("La Cueva",
                                 "La Caoba",
                                 "Morelia",
                                 "El Corral",
                                 "Aceitillar"))
```


## Cast for site persistent

```{r}
#sample size
sp.cast.n <- dcast(data= site_persist.foc[,],
      formula = spp.code + stat ~ site,
      value.var = "site",
      fun.aggregate = length)

#siter persisten individuals
sp.cast.persist <- dcast(data= site_persist.foc[,],
      formula = spp.code + stat ~ site+persist,
      value.var = "site",
      fun.aggregate = length)


```


```{r}
#isolate focal columns
j <- grep("_P",names(sp.cast.persist))
sp.cast.persist[,j]

sp.cast.persist2 <- round(sp.cast.persist[,j]/sp.cast.n[,-c(1:2)],2)


```

### Paste together 


```{r}
sp.cast.persist3 <- sp.cast.persist2

sp.cast.persist3$`La Cueva_P` <- paste(sp.cast.persist3$`La Cueva_P`," (",sp.cast.n$`La Cueva`,")", sep = "")

sp.cast.persist3$`La Caoba_P` <- paste(sp.cast.persist3$`La Caoba_P`," (",sp.cast.n$`La Caoba`,")", sep = "")

sp.cast.persist3$Morelia_P <- paste(sp.cast.persist3$`Morelia_P`," (",sp.cast.n$Morelia,")", sep = "")

sp.cast.persist3$`El Corral_P` <- paste(sp.cast.persist3$`El Corral_P`," (",sp.cast.n$`El Corral`,")", sep = "")
```

```{r}
table5_6_site_persist <- data.frame(spp.code = sp.cast.persist[,1],
                                    sp.cast.persist3)
```



### Save table

```{r}
write.csv(table5_6_site_persist, 
          file = "./tables/temp_tables_5_6_site_persist.csv",
          row.names = F)
```

