---
title: "Make Tables 5 & 6: sample size, age and sex"
author: "Nathan Brouwer"
date: "January 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script create the Sampe size, age ratio, and sex ratio part of tables 5 and 6
It uses a dataset where Mencia and Aceitillar data have been combined into a single long-format dataframe

## Libraries

```{r}
library(reshape2)
library(here)

```


## Load data

Load annual Captures data with Mencia AND Aceitillar combined


```{r}
filename <- "annual_captures.RData"
load(here::here("data",filename))
```



## Subset just by focal species

These are the spp focused on in teh original MS; they are the most commonly caught


Names of species
```{r}
focal.mig <- c("OVEN","BAWW","COYE","AMRE","CMWA",
               "BTBW","PAWA","PRAW")

focal.res <- c("HILC" # formerly HLCU
,"STOF"
,"RLTH"
,"NOMO"
,"GTGT"   # formerly GRWA
,"BANA"
,"BCPT"
,"YFGR"
,"BFGR"
,"GABU")

focals <- c(focal.mig,focal.res)
```

Subset
```{r}
i.focals <- which(ann_caps$spp.code %in% focals)

ann_caps.foc <- ann_caps[i.focals,]
dim(ann_caps.foc)
dim(ann_caps)
```

## Set spp order for tables


replicate order from  tables in oridingal MS

```{r}
ann_caps.foc$spp.code <- factor(ann_caps.foc$spp.code,
                             levels = focals)
```


## Order sites by age

```{r}
ann_caps.foc$site <- factor(ann_caps.foc$site,
                      levels = c("La Cueva",
                                 "La Caoba",
                                 "Morelia",
                                 "El Corral",
                                 "Aceitillar"))
```




### Create Unique ID

Index of that yields each individual only once per year, but allows individuals to appear in subsequent years.

dataframe has been organized by date so this should select the 1st time an individual was captured

```{r}
i.within.yr  <- match(unique(ann_caps.foc$ID.yr),
                      ann_caps.foc$ID.yr)
```


### Calcualte sample size

```{r}
cast.N <- dcast(data= ann_caps.foc[i.within.yr,],
      formula = spp.code + status2 ~ site,
      value.var = "site",
      fun.aggregate = length)
```


### cast for sex ratio
 

```{r}
cast.sex <- dcast(data= ann_caps.foc[i.within.yr,],
      formula = spp.code + status2 ~ site + sex.M,
      value.var = "site",
      fun.aggregate = length)
```

### cast for  age
```{r}
cast.age <- dcast(data= ann_caps.foc[i.within.yr,],
      formula = spp.code + status2 ~ site + age.AHY.01,
      value.var = "site",
      fun.aggregate = length)
```


### Calcualte sex ratios

```{r}
la.cueva.M <- cast.sex$`La Cueva_M`/(cast.sex$`La Cueva_M`+cast.sex$`La Cueva_other`)
la.cueva.M.n <- cast.sex$`La Cueva_M`+cast.sex$`La Cueva_other`


la.caoba.M <- cast.sex$`La Caoba_M`/(cast.sex$`La Caoba_M`+cast.sex$`La Caoba_other`)
la.caoba.M.n <- cast.sex$`La Caoba_M`+cast.sex$`La Caoba_other`
  
morelia.M <- cast.sex$`Morelia_M`/(cast.sex$`Morelia_M`+cast.sex$`Morelia_other`)
morelia.M.n <- cast.sex$`Morelia_M`+cast.sex$`Morelia_other`
  
  
corral.M   <- cast.sex$`El Corral_M`/(cast.sex$`El Corral_M`+ cast.sex$`El Corral_other`)
Corral_M.n <-                         cast.sex$`El Corral_M`+ cast.sex$`El Corral_other`

Aceitillar.M <- cast.sex$`Aceitillar_M`/(cast.sex$`Aceitillar_M`+cast.sex$`Aceitillar_other`)
Aceitillar_M.n <- cast.sex$`Aceitillar_M`+cast.sex$`Aceitillar_other`
```





### Calcualte age ratios

```{r}
la.cueva.AHY <- cast.age$`La Cueva_age1`/(cast.age$`La Cueva_age1`+cast.age$`La Cueva_age2`)
la.cueva.AHY.n <- cast.age$`La Cueva_age1`+cast.age$`La Cueva_age2`
  

la.caoba.AHY <- cast.age$`La Caoba_age1`/(cast.age$`La Caoba_age1`+cast.age$`La Caoba_age2`)
la.caoba.AHY.n <- cast.age$`La Caoba_age1`+cast.age$`La Caoba_age2`


morelia.AHY <- cast.age$`Morelia_age1`/(cast.age$`Morelia_age1`+cast.age$`Morelia_age2`)
morelia.AHY.n <- cast.age$`Morelia_age1`+cast.age$`Morelia_age2`
  

corral.AHY   <- cast.age$`El Corral_age1`/(cast.age$`El Corral_age1`+cast.age$`El Corral_age2`)
corral.AHY.n <-                            cast.age$`El Corral_age1`+cast.age$`El Corral_age2`


Aceitillar.AHY <- cast.age$`Aceitillar_age1`/(cast.age$`Aceitillar_age1`+cast.age$`Aceitillar_age2`)
Aceitillar.AHY.n <- cast.age$`Aceitillar_age1`+cast.age$`Aceitillar_age2`
```




### Build table

```{r}
dig <- 2
tables5.6 <- data.frame(cast.N[,c("spp.code","status2")],
           La.Cueva.N =  cast.N$`La Cueva`,      
           La.Cueva.M = round(la.cueva.M,2),
           la.cueva.M.n = la.cueva.M.n,
           La.Cueva.AHY =   round(la.cueva.AHY,2),
           la.cueva.AHY.n = la.cueva.AHY.n,
           
           La.Caoba.N = cast.N$`La Caoba`,
           La.Caoba.M =  round(la.caoba.M,2),
           la.caoba.M.n = la.caoba.M.n,
           La.Caoba.AHY =   round(la.caoba.AHY,2),
           la.caoba.AHY.n = la.caoba.AHY.n,
           
           Morelia.N = cast.N$Morelia,
           Morelia.M =   round(morelia.M,2),
           morelia.M.n = morelia.M.n,
           Morelia.AHY = round(morelia.AHY,2),
           morelia.AHY.n = morelia.AHY.n,
           
           El.Corral.N = cast.N$`El Corral`,
           El.Corral.M =  round(corral.M,2),
           Corral_M.n = Corral_M.n,
           El.Corral.AHY =    round(corral.AHY,2),
           corral.AHY.n = corral.AHY.n,
           
           Aceitillar.N = cast.N$`Aceitillar`,
           Aceitillar.M =  round(Aceitillar.M,2),
           Aceitillar_M.n = Aceitillar_M.n,
           Aceitillar.AHY =    round(Aceitillar.AHY,2),
           Aceitillar.AHY.n = Aceitillar.AHY.n)
```




### Paste together 

```{r}
make.column <- function(df,col.name,sample.size){
  x <- paste(df[,col.name]," (",df[,sample.size],")", sep = "")
  x[grep("N",x)] <- ""
  return(x)
}
```


```{r}
La.Cueva.M <- make.column(tables5.6,
                          "La.Cueva.M","la.cueva.M.n")
La.Caoba.M <- make.column(tables5.6,
                          "La.Caoba.M","la.caoba.M.n")
Morelia.M <- make.column(tables5.6,
                          "Morelia.M","morelia.M.n")
El.Corral.M <- make.column(tables5.6,
                          "El.Corral.M","Corral_M.n")

Aceitillar.M <- make.column(tables5.6,
                          "Aceitillar.M","Aceitillar_M.n")
```


```{r}
La.Cueva.AHY <- make.column(tables5.6,
                          "La.Cueva.AHY","la.cueva.AHY.n")
La.Caoba.AHY <- make.column(tables5.6,
                          "La.Caoba.AHY","la.caoba.AHY.n")
Morelia.AHY <- make.column(tables5.6,
                          "Morelia.AHY","morelia.AHY.n")
El.Corral.AHY <- make.column(tables5.6,
                          "El.Corral.AHY","corral.AHY.n")

Aceitillar.AHY <- make.column(tables5.6,
                          "Aceitillar.AHY","Aceitillar.AHY.n")


```

```{r}
data.frame(El.Corral.AHY,
tables5.6$El.Corral.AHY,
tables5.6$Corral_M.n)
```


```{r}
tables5.6.final <- data.frame(tables5.6[,c("spp.code","status2")],
           La.Cueva.N = tables5.6$La.Cueva.N,
            La.Cueva.AHY,
            La.Cueva.M,
           
           La.Caoba.N = tables5.6$La.Caoba.N,
            La.Caoba.AHY,
            La.Caoba.M,
           
           Morelia.N = tables5.6$Morelia.N,
            Morelia.AHY,
            Morelia.M,
           
           El.Corral.N = tables5.6$El.Corral.N,
            El.Corral.AHY,
            El.Corral.M,
           
           Aceitillar.N = tables5.6$Aceitillar.N,
            Aceitillar.AHY,
            Aceitillar.M)
```



### Save table

```{r}
write.csv(tables5.6.final, 
          file = "./tables/temp_tables_5_6_age_etc.csv",
          row.names = F)
```

```{r}

```

