---
title: "Clean insect data taken from leaf litter samples"
author: "Nathan Brouwer"
date: "February 7, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Clean and tidy data taken from leaf litter samples.

## Preliminaries

### Load libraries

```{r}
library(here)
library(reshape2)
library(stringr)
library(vegan)
library(mvabund)
library(ggplot2)
library(cowplot)
library(lme4)
library(blme)
library(coefplot)
library("rptR")
```


### Load data

#### Construct file path
```{r}
file. <- here::here("data-raw",
           "data-raw-insects",
           "data-raw-leaf-litter",
           "CSV_leaf_litter_both_sites",
           "leaf_litter_insects_CSV.csv")
```

#### load data
```{r}
litter <- read.csv(file= file.,
                   skip = 6,
                   header = TRUE,
                   blank.lines.skip = TRUE,
                   na.strings = "")
```

#### Check wide data
```{r}
explore <- TRUE
if(explore  == TRUE){
  head(litter)
  
  summary(litter,15)
  
  dim(litter)
}
```


## Clean wide data

### Remove blank rows

```{r}
litter <- na.omit(litter)
```

### Change column names
```{r}
names(litter) <- gsub("^X","S",names(litter))
```


### Wide data structure

```{r eval=FALSE, include=FALSE}
head(litter)
```

       site month      taxa S1 S2 S3 S4 S5 S6 S7 S8 S9 S10 S11 S12 S13 S14 S15 S16 S17 S18
1 Broadleaf   Jan   Spiders  0  0  0  2  3  2  1  1  0   1   2   1   6   1   0   2   0   0
2 Broadleaf   Jan      Ants  0  0  0  4  1  1  3  0  0   0   1   0   5   0   0   0   0   0
3 Broadleaf   Jan Cockroach  0  1  0  0  1  2  1  0  0   2   1   0   5   1   0   0   0   1
4 Broadleaf   Jan    Beetle  0  0  0  0  0  0  0  0  0   0   0   1   0   0   0   0   0   0
5 Broadleaf   Jan   Cricket  0  0  0  2  0  0  0  0  0   0   0   0   0   0   0   0   0   0
6 Broadleaf   Jan    Gusano  0  0  0  0  0  0  0  0  0   0   0   0   1   1   0   0   0   0
  
## Melt data

### Melt

Melt data from wide to long.
```{r}
id.vars. <- names(litter)[-grep("^S",names(litter))]
litter.melt <- melt(data = litter,
                    id.vars = id.vars.,
                    variable.name = "sample",
                    value.name = "N")
                    
```

### Clean melted data

```{r}
summary(factor(litter.melt$sample))
```

```{r}

litter.melt$sample <- gsub("S","",litter.melt$sample )
summary(factor(litter.melt$sample))
litter.melt$sample <- as.numeric(litter.melt$sample)
```


#### Check cleaned wide data
```{r}
explore <- TRUE
wrking <- litter.melt
if(explore  == TRUE){
  head(wrking)
  
  summary(wrking,15)
  
  dim(wrking)
}
```













### Plot

```{r}
litter.melt$site <- factor(litter.melt$site,
                           levels = c("La Cueva","La Caoba","Morelia","El Corral",
                                      "Broadleaf"))
```



```{r}
ggplot(data = litter.melt,
       aes(y = N,
           x = site,
           color = site)) +
  geom_boxplot() +
facet_wrap(~ taxa,scale = "free")
```



### ggplot

```{r}
 #ggplot2 time saver: calculate mean SE/CIs on the fly
ggplot(data = litter.melt,
       aes(y = N,
           x = site,
           color = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```




```{r}
library(ggpubr)

ggerrorplot(data = litter.melt,
            y = "N",
            x = "site",
            color = "site",
            desc_stat = "mean_ci",
            size = 2)
```




## Save Long data

```{r}
litter <- litter.melt
save(litter, file = "./data/litter.RData")
```

