---
title: "Clean insect data taken from leaf flying samples"
author: "Nathan Brouwer"
date: "February 7, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Clean and tidy data taken from leaf flying samples.

## Preliminaries

### Load libraries

```{r}
library(here)
library(reshape2)
library(stringr)
library(vegan)
library(mvabund)
library(ggplot2)
library(cowplot)
library(lme4)
library(blme)
library(coefplot)
library("rptR")
```


### Load data

#### Construct file path
```{r}
file. <- here::here("data-raw",
           "data-raw-insects",
           "data-raw-sticky-traps",
           "CSV_sticky_traps_both_sites",
           "CSV_stick_traps_both_sites.csv")
```

#### load data
```{r}
flying <- read.csv(file= file.,
                   skip = 6,
                   header = TRUE,
                   blank.lines.skip = TRUE,
                   na.strings = c("","NA"))
```

#### Check wide data
```{r}
explore <- TRUE
if(explore  == TRUE){
  head(flying)
  
  summary(flying,15)
  
  dim(flying)
}
```


## Clean wide data

### Change column names
```{r}
names(flying) <- gsub("^X","S",names(flying))
```


### Wide data structure

```{r eval=FALSE, include=FALSE}
head(flying)
```


### Change 

       site        taxa   S1  S2   S3  S4   S5   S6   S7   S8   S9  S10 S11 S12
1 Broadleaf  Coleoptera  1.0 0.5  3.5 0.0  2.0  1.5  8.0  0.0  0.0  4.5 1.5  31
2 Broadleaf     Diptera  4.0 5.5 15.0 5.5  9.5 13.5 12.0 14.0 22.5 10.5  19  22
3 Broadleaf   Homoptera 37.0 5.5  7.5 7.0 17.5 22.5 22.5 12.0 29.5 18.0 6.5  11
4 Broadleaf Hymenoptera  2.5 1.0  3.0 0.5  0.0  4.5  6.0  2.5  3.5  2.5   4   8
5 Broadleaf Lepidoptera  0.0 0.0  0.0 0.0  0.0  0.0  0.0  0.0  0.0  0.0   0   0
6 Broadleaf     Spiders  0.0 0.5  0.0 0.0  0.5  0.0  0.0  0.0  0.5  0.0   0 0.5

## Melt data

### Melt

Melt data from wide to long.
```{r}
id.vars. <- names(flying)[-grep("^S",names(flying))]

flying.melt <- melt(data = flying[,],
                    id.vars = id.vars.,
                    variable.name = "sample",
                    value.name = "N",factorsAsStrings = TRUE)



summary(flying.melt)                     
```

### Clean melted data

```{r}
summary(factor(flying.melt$sample))
```

```{r}

flying.melt$sample <- gsub("S","",flying.melt$sample )
summary(factor(flying.melt$sample))
flying.melt$sample <- as.numeric(flying.melt$sample)
```


#### Check cleaned wide data
```{r}
explore <- TRUE
wrking <- flying.melt
if(explore  == TRUE){
  head(wrking)
  
  summary(wrking,15)
  
  dim(wrking)
}
```













### Set factor levels

```{r}
flying.melt$site <- factor(flying.melt$site,
                           levels = c("La Cueva","La Caoba","Morelia","El Corral",
                                      "Broadleaf"))
```


## Plot 
```{r}
ggplot(data = flying.melt,
       aes(y = N,
           x = site,
           color = site)) +
  geom_boxplot() +
facet_wrap(~ taxa,scale = "free")
```



### ggplot

```{r}
 #ggplot2 time saver: calculate mean SE/CIs on the fly
ggplot(data = flying.melt,
       aes(y = N,
           x = site,
           color = site)) +
  stat_summary(fun.data  = "mean_cl_boot",size = 2)
```




```{r}
library(ggpubr)

ggerrorplot(data = flying.melt,
            y = "N",
            x = "site",
            color = "site",
            desc_stat = "mean_ci",
            size = 2)
```




## Save Long data

```{r}
flying <- flying.melt
save(flying, file = "./data/flying.RData")
```

