---
title: "1b_reshape_mencia_site_persistence"
author: "Nathan Brouwer"
date: "January 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Libraries
```{r}
library(reshape2)
library(stringr)
library(stringi)
library(lubridate)
library(here)
```



### File loading function
clunky but works
```{r}
filename <- here::here("data-raw",
                  "data-raw-mencia",
                  "mencia_site_persistence",
                 "mencia_site_persist_cleaned")
print(filename)

skip <- 0
sp.wide <- read.csv(file = filename,
                skip = skip,
                header =TRUE,
                strip.white = TRUE,
                blank.lines.skip = TRUE,
                na.strings = c("NA",""," "),
                stringsAsFactors = FALSE,
                colClasses=c("SUF"="character")
                )


```



## Shorten names

```{r}
names(sp.wide) <- gsub("resight","rs",names(sp.wide))
names(sp.wide) <- gsub("recap","rc",names(sp.wide))
names(sp.wide) <- gsub("20","",names(sp.wide))
names(sp.wide) <- gsub("nov","N",names(sp.wide))
names(sp.wide) <- gsub("jan","J",names(sp.wide))
names(sp.wide) <- gsub("feb","F",names(sp.wide))
names(sp.wide) <- gsub("wanderer","wander",names(sp.wide))
names(sp.wide) <- gsub("wander","wndr",names(sp.wide))
names(sp.wide) <- gsub("species","spp",names(sp.wide))
names(sp.wide) <- gsub("wing","wng",names(sp.wide))
names(sp.wide) <- gsub("tarsus","tar",names(sp.wide))
names(sp.wide) <- gsub("year","yr",names(sp.wide))
names(sp.wide) <- gsub("month","mo",names(sp.wide))
names(sp.wide) <- gsub("band","bnd",names(sp.wide))
names(sp.wide) <- gsub("sex","sx",names(sp.wide))
names(sp.wide) <- gsub("colors","col",names(sp.wide))
names(sp.wide) <- gsub("notes","nts",names(sp.wide))
names(sp.wide) <- gsub("status","stat",names(sp.wide))
names(sp.wide) <- gsub("age","ag",names(sp.wide))
```

## Change "n/a" to "bf" for "before" 1st capture

Steve has "n/a" list for sampling occassions w/in a season before a bird was 1st observe (but not for years b/f)/

Change these to "bf"

```{r}
natobf<- function(x){ifelse(x == "n/a","bf",x)}
j.r <- grep("^r", names(sp.wide))

sp.wide[,j.r] <- apply(sp.wide[,j.r],2, natobf)
```



## Change "X" to 1

steve list "X" for an observation or recapture.  
change to 1 to be consistent the mark recaput

```{r}
Xto1<- function(x){ifelse(x == "X" | x == "x","1",x)}
j.r <- grep("^r", names(sp.wide))

sp.wide[,j.r] <- apply(sp.wide[,j.r],2, Xto1)
```

## Change site names to 1

steve list site names in some cases instead of "X".

Change to 1

```{r}
siteto1<- function(x){
  i <- which(x %in% c("Corral","Caoba","Cueva","Morelia","Cue/Cao",
                      "Fuera","DIED",
                      "repeated combo"))
  x[i] <- 1
  return(x)
  }

sp.wide[,j.r] <- apply(sp.wide[,j.r],2, siteto1)
```

## Convert all to factor
```{r}
for(i in 1:length(j.r)){
  j <- j.r[i]
  sp.wide[,j] <- as.factor(sp.wide[,j])
}
summary(sp.wide[,j.r])
summary(factor(sp.wide$rs.04.J))
```


## Calculate number of times captured

```{r}
calc_caps <- function(x){
  return(sum(as.numeric(x[grep("1",x)])))
}

sp.wide$num.rc.rs <- apply(sp.wide[,j.r],1, calc_caps)
```



```{r}
i.drop1 <- which(names(sp.wide) %in% c("bnd.pre",#"bnd.suf",
                                       "wng","tar",
                                       "site.ag.init",
                                       "site.ag","bnd",
                                       "nts","yr.num",
                                       "date2","mo","day"))
                 
sp.wide[,-c(j.r,i.drop1)]
```


## Were any wanders incorrectly coded?

None

```{r}
which(sp.wide$wndr == "W" & sp.wide$num.rc.rs > 0)
```

## Were any P (=persistent) never resighted or recaptured?

nope 

```{r}
which(sp.wide$wndr == "P" & sp.wide$num.rc.rs < 1)


```


## Are all residents coded 

Yes

```{r}
i.prob <- which(sp.wide$wndr == "P" & sp.wide$num.rc.rs < 1)

sp.wide[i.prob, -c(j.r,i.drop1)]
```




## Examine wanderer

```{r}
summary(factor(sp.wide$wndr ))

i.w <- which(sp.wide$wndr == "W")
j.r <- grep("^r", names(sp.wide))

sp.wide[i.w, j.r]
```


## Code "P"

```{r}
sp.wide$persist <- ifelse(sp.wide$num.rc.rs > 0, "P","NP")
```




## Change spp to spp.code

```{r}
names(sp.wide) <- gsub("spp","spp.code",names(sp.wide))
```


## Load spp meta / trait info

```{r}
spp.meta <- read.csv(here::here("data","spp_list.csv"),stringsAsFactors =F)
```


Subset columns I want to use
```{r}
spp.meta <- spp.meta[c("spp.code","spp",
                       "status2",
                       "hab1","diet")]


```

```{r}
summary(spp.meta)
```

## Change GRWA -> GTGT

Green tailed ground warlber and green tailed ground tanager are the same

Change column name

```{r}
names(sp.wide) <- gsub("species","spp.code", names(sp.wide))
```


```{r}
sp.wide$spp.code <- as.character(sp.wide$spp.code)
sp.wide$spp.code[which(sp.wide$spp.code == "GRWA")] <- "GTGT"
```

```{r}
sp.wide$spp.code[which(sp.wide$spp.code == "HLCU")] <- "HILC"

```


MYWA =  myrtyle warbler = YRWA yellow rumped warbler
NUMA -> SBMU #nutmet manakin = scaly breasted munia
WHQD = WFQD

```{r}
sp.wide$spp.code[which(sp.wide$spp.code == "MYWA")] <- "YRWA"
sp.wide$spp.code[which(sp.wide$spp.code == "NUMA")] <- "SBMU"
sp.wide$spp.code[which(sp.wide$spp.code == "WHQD")] <- "WFQD"
```


## Merge cleaned data w/ species meta data


Merge
```{r}

names(sp.wide)
names(spp.meta)
names(sp.wide2)

sp.wide2 <- merge(sp.wide, spp.meta, all = T)

dim(sp.wide)
dim(sp.wide2)

sp.wide2 <-sp.wide2[-which(is.na(sp.wide2$site) == TRUE), ]
sp.wide <- sp.wide2
```


